---
title: "Data Collection"
date: "`r Sys.Date()`"
format: html
execute:
  echo: true
---

### ============================================

### Data Collection Script

### This script collects and processes raw data from various sources

### Sources include:

### - 1. Retrieve USDA NASS pastureland data from USDA Census of Agriculture

### - 2. Retrieve USDA NASS cattle inventory data from USDA Census of Agriculture

### - 3. Retrieve hay production data from USDA NASS 

### - 4. Retrieve cattle dressing weights data from USDA NASS

### - 5. Retrieve weather data from PRISM

### - 6. Retrieve Climate projections data 

### - 7.Retrieve USDM data

###  - 8.Retrieve PSDI data


### =============================================

```{r}
# Load packages
pacman::p_load(
  stars,        # spatiotemporal data handling
  terra,        # raster data handling
  sf,           # vector data handling
  dplyr,        # data wrangling
  tidyverse,
  fixest,
  usdarnass,
  flextable,
  tigris,
  zoo,
  purrr,
  ncdf4,
  stringr,      # string manipulation
  lubridate,    # date handling
  data.table,   # fast data wrangling
  tidyr,        # reshape
  keyring,      # API key management
  ggplot2,
  tmap,
  future.apply, # parallel processing
  prism,        # download PRISM data
  exactextractr,
  tidycensus,
  readxl,
  readr,
  patchwork
)
rm(list = ls())

```

### 1. Retrieve USDA NASS pastureland data from USDA Census of Agriculture

```{r}
#| echo: false

clean_numeric <- function(x) suppressWarnings(readr::parse_number(as.character(x)))
pasture_after1997 <- readxl::read_xlsx("/Users/sharaakat/Dropbox/akat_shara/DWFI_drought/GitHub/Data/Data/pasture_data_after1997.xlsx") %>%
  dplyr::select(State, County, Year, Value) %>%
  rename(state = State, county = County, year = Year, pastureland_all_types = Value) %>%
  mutate(across(c(year, pastureland_all_types), clean_numeric),  # <-- this was missing
         state  = tolower(state),
         county = tolower(county))

# County pastureland (pre-1997, hand-digitized from PDFs)
pasture_before1997 <- readxl::read_xlsx("/Users/sharaakat/Dropbox/akat_shara/DWFI_drought/GitHub/Data/Data/pasture_data_before1997.xlsx") %>%
  mutate(across(c(year, pastureland_all_types), clean_numeric),
         state  = tolower(state),
         county = tolower(county))

# Combine
pasture_data_ALL <- bind_rows(pasture_before1997, pasture_after1997)


```

### 2. Retrieve USDA NASS cattle inventory data from USDA Census of Agriculture

```{r}

# National inventory (post-1997)
clean_numeric <- function(x) suppressWarnings(readr::parse_number(as.character(x)))
beef_inv_national <- readxl::read_xlsx("/Users/sharaakat/Dropbox/akat_shara/DWFI_drought/GitHub/Data/Data/county_beef_after1997_National.xlsx") %>%
  dplyr::select(Year, Value) %>%
  rename(year = Year, beef_inv = Value) %>%
  mutate(across(everything(), clean_numeric))

# County inventory (post-1997)
beef_inv_after1997 <- readxl::read_xlsx("/Users/sharaakat/Dropbox/akat_shara/DWFI_drought/GitHub/Data/Data/county_beefinv_after1997.xlsx") %>%
  dplyr::select(State, County, Year, Value) %>%
  rename(state = State, county = County, year = Year, beef_inv = Value) %>%
  mutate(across(c(year, beef_inv), clean_numeric),
         state  = tolower(state),
         county = tolower(county))

# County inventory (pre-1997, hand-digitized from PDFs)
beef_inv_before1997 <- readxl::read_xlsx("/Users/sharaakat/Dropbox/akat_shara/DWFI_drought/GitHub/Data/Data/county_beefinv_before1997.xlsx") %>%
  mutate(across(c(year, beef_inv), clean_numeric),
         state  = tolower(state),
         county = tolower(county))

# Combine
beef_inv <- bind_rows(beef_inv_before1997, beef_inv_after1997)

```

### 3. Retrieve hay production data from USDA NASS 

```{r}

# State abbreviation to full name lookup
state_names <- c(
  AL = "Alabama",       AK = "Alaska",        AZ = "Arizona",       AR = "Arkansas",
  CA = "California",    CO = "Colorado",       CT = "Connecticut",   DE = "Delaware",
  FL = "Florida",       GA = "Georgia",        HI = "Hawaii",        ID = "Idaho",
  IL = "Illinois",      IN = "Indiana",        IA = "Iowa",          KS = "Kansas",
  KY = "Kentucky",      LA = "Louisiana",      ME = "Maine",         MD = "Maryland",
  MA = "Massachusetts", MI = "Michigan",       MN = "Minnesota",     MS = "Mississippi",
  MO = "Missouri",      MT = "Montana",        NE = "Nebraska",      NV = "Nevada",
  NH = "New Hampshire", NJ = "New Jersey",     NM = "New Mexico",    NY = "New York",
  NC = "North Carolina",ND = "North Dakota",   OH = "Ohio",          OK = "Oklahoma",
  OR = "Oregon",        PA = "Pennsylvania",   RI = "Rhode Island",  SC = "South Carolina",
  SD = "South Dakota",  TN = "Tennessee",      TX = "Texas",         UT = "Utah",
  VT = "Vermont",       VA = "Virginia",       WA = "Washington",    WV = "West Virginia",
  WI = "Wisconsin",     WY = "Wyoming"
)

# State-level hay stocks (1,000 tons)
hay_all_stocks_1000t <- readxl::read_xlsx(
  "/Users/sharaakat/Dropbox/akat_shara/DWFI_drought/GitHub/RobustnessCheck/Data/StateHaySupplies.xlsx",
  sheet = 2, skip = 2
) %>%
  pivot_longer(cols = -year, names_to = "state", values_to = "hay_stock_1000t") %>%
  mutate(state = tolower(state_names[state])) %>%
  group_by(state) %>%
  mutate(
    # Pre-1986: reported as of Jan 1 → lead by 1 year to align with Dec 1 convention
    # Post-1986: reported as of Dec 1 → already beginning-of-year inventory, no shift needed
    hay_stock_1000t = ifelse(year < 1986, dplyr::lead(hay_stock_1000t, n = 1), hay_stock_1000t)
  ) %>%
  ungroup()

# National average hay stocks by year
hay_all_stocks_1000t_national <- hay_all_stocks_1000t %>%
  group_by(year) %>%
  summarize(hay_stock_1000t = mean(hay_stock_1000t, na.rm = TRUE))

# HAY STOCKS average for neighbouring states                           
neighbors_raw <- tribble(
  ~state_name, ~state, ~nbr_name, ~nbr,
  "Alabama","AL","Florida","FL",
  "Alabama","AL","Georgia","GA",
  "Alabama","AL","Mississippi","MS",
  "Alabama","AL","Tennessee","TN",
  "Alaska","AK","","",
  "Arizona","AZ","California","CA",
  "Arizona","AZ","Colorado","CO",
  "Arizona","AZ","Nevada","NV",
  "Arizona","AZ","New Mexico","NM",
  "Arizona","AZ","Utah","UT",
  "Arkansas","AR","Louisiana","LA",
'Arkansas','AR','Mississippi','MS',
'Arkansas','AR','Missouri','MO',
'Arkansas','AR','Oklahoma','OK',
'Arkansas','AR','Tennessee','TN',
'Arkansas','AR','Texas','TX',
'California','CA','Arizona','AZ',
'California','CA','Nevada','NV',
'California','CA','Oregon','OR',
'Colorado','CO','Arizona','AZ',
'Colorado','CO','Kansas','KS',
'Colorado','CO','Nebraska','NE',
'Colorado','CO','New Mexico','NM',
'Colorado','CO','Oklahoma','OK',
'Colorado','CO','Utah','UT',
'Colorado','CO','Wyoming','WY',
'Connecticut','CT','Massachusetts','MA',
'Connecticut','CT','New York','NY',
'Connecticut','CT','Rhode Island','RI',
'Delaware','DE','Maryland','MD',
'Delaware','DE','New Jersey','NJ',
'Delaware','DE','Pennsylvania','PA',
'Florida','FL','Alabama','AL',
'Florida','FL','Georgia','GA',
'Georgia','GA','Alabama','AL',
'Georgia','GA','Florida','FL',
'Georgia','GA','North Carolina','NC',
'Georgia','GA','South Carolina','SC',
'Georgia','GA','Tennessee','TN',
'Hawaii','HI','','',
'Idaho','ID','Montana','MT',
'Idaho','ID','Nevada','NV',
'Idaho','ID','Oregon','OR',
'Idaho','ID','Utah','UT',
'Idaho','ID','Washington','WA',
'Idaho','ID','Wyoming','WY',
'Illinois','IL','Indiana','IN',
'Illinois','IL','Iowa','IA',
'Illinois','IL','Michigan','MI',
'Illinois','IL','Kentucky','KY',
'Illinois','IL','Missouri','MO',
'Illinois','IL','Wisconsin','WI',
'Indiana','IN','Illinois','IL',
'Indiana','IN','Kentucky','KY',
'Indiana','IN','Michigan','MI',
'Indiana','IN','Ohio','OH',
'Iowa','IA','Illinois','IL',
'Iowa','IA','Minnesota','MN',
'Iowa','IA','Missouri','MO',
'Iowa','IA','Nebraska','NE',
'Iowa','IA','South Dakota','SD',
'Iowa','IA','Wisconsin','WI',
'Kansas','KS','Colorado','Co',
'Kansas','KS','Missouri','MO',
'Kansas','KS','Nebraska','NE',
'Kansas','KS','Oklahoma','OK',
'Kentucky','KY','Illinois','IL',
'Kentucky','KY','Indiana','IN',
'Kentucky','KY','Missouri','MO',
'Kentucky','KY','Ohio','OH',
'Kentucky','KY','Tennessee','TN',
'Kentucky','KY','Virginia','VA',
'Kentucky','KY','West Virginia','WV',
'Louisiana','LA','Arkansas','AR',
'Louisiana','LA','Mississippi','MS',
'Louisiana','LA','Texas','TX',
'Maine','ME','New Hampshire','NH',
'Maryland','MD','Delaware','DE',
'Maryland','MD','Pennsylvania','PA',
'Maryland','MD','Virginia','VA',
'Maryland','MD','West Virginia','WV',
'Massachusetts','MA','Connecticut','CT',
'Massachusetts','MA','New Hampshire','NH',
'Massachusetts','MA','New York','NY',
'Massachusetts','MA','Rhode Island','RI',
'Massachusetts','MA','Vermont','VT',
'Michigan','MI','Illinois','IL',
'Michigan','MI','Indiana','IN',
'Michigan','MI','Minnesota','MN',
'Michigan','MI','Ohio','OH',
'Michigan','MI','Wisconsin','WI',
'Minnesota','MN','Iowa','IA',
'Minnesota','MN','Michigan','MI',
'Minnesota','MN','North Dakota','ND',
'Minnesota','MN','South Dakota','SD',
'Minnesota','MN','Wisconsin','WI',
'Mississippi','MS','Alabama','AL',
'Mississippi','MS','Arkansas','AR',
'Mississippi','MS','Louisiana','LA',
'Mississippi','MS','Tennessee','TN',
'Missouri','MO','Arkansas','AR',
'Missouri','MO','Illinois','IL',
'Missouri','MO','Iowa','IA',
'Missouri','MO','Kansas','KS',
'Missouri','MO','Kentucky','KY',
'Missouri','MO','Nebraska','NE',
'Missouri','MO','Oklahoma','OK',
'Missouri','MO','Tennessee','TN',
'Montana','MT','Idaho','ID',
'Montana','MT','North Dakota','ND',
'Montana','MT','South Dakota','SD',
'Montana','MT','Wyoming','WY',
'Nebraska','NE','Colorado','CO',
'Nebraska','NE','Iowa','IA',
'Nebraska','NE','Kansas','KS',
'Nebraska','NE','Missouri','MO',
'Nebraska','NE','South Dakota','SD',
'Nebraska','NE','Wyoming','WY',
'Nevada','NV','Arizona','AZ',
'Nevada','NV','California','CA',
'Nevada','NV','Idaho','ID',
'Nevada','NV','Oregon','OR',
'Nevada','NV','Utah','UT',
'New Hampshire','NH','Maine','ME',
'New Hampshire','NH','Massachusetts','MA',
'New Hampshire','NH','Vermont','VT',
'New Jersey','NJ','Delaware','DE',
'New Jersey','NJ','New York','NY',
'New Jersey','NJ','Pennsylvania','PA',
'New Mexico','NM','Arizona','AZ',
'New Mexico','NM','Colorado','CO',
'New Mexico','NM','Oklahoma','OK',
'New Mexico','NM','Texas','TX',
'New Mexico','NM','Utah','UT',
'New York','NY','Connecticut','CT',
'New York','NY','Massachusetts','MA',
'New York','NY','New Jersey','NJ',
'New York','NY','Pennsylvania','PA',
'New York','NY','Rhode Island','RI',
'New York','NY','Vermont','VT',
'North Carolina','NC','Georgia','GA',
'North Carolina','NC','South Carolina','SC',
'North Carolina','NC','Tennessee','TN',
'North Carolina','NC','Virginia','VA',
'North Dakota','ND','Minnesota','MN',
'North Dakota','ND','Montana','MT',
'North Dakota','ND','South Dakota','SD',
'Ohio','OH','Indiana','IN',
'Ohio','OH','Kentucky','KY',
'Ohio','OH','Michigan','MI',
'Ohio','OH','Pennsylvania','PA',
'Ohio','OH','West Virginia','WV',
'Oklahoma','OK','Arkansas','AR',
'Oklahoma','OK','Colorado','CO',
'Oklahoma','OK','Kansas','KS',
'Oklahoma','OK','Missouri','MO',
'Oklahoma','OK','New Mexico','NM',
'Oklahoma','OK','Texas','TX', 
'Oregon','OR','California','CA',
'Oregon','OR','Idaho','ID',
'Oregon','OR','Nevada','NV',
'Oregon','OR','Washington','WA',
'Pennsylvania','PA','Delaware','DE',
'Pennsylvania','PA','Maryland','MD',
'Pennsylvania','PA','New Jersey','NJ',
'Pennsylvania','PA','New York','NY',
'Pennsylvania','PA','Ohio','OH',
'Pennsylvania','PA','West Virginia','WV',
'Rhode Island','RI','Connecticut','CT',
'Rhode Island','RI','Massachusetts','MA',
'Rhode Island','RI','New York','NY',
'South Carolina','SC','Georgia','GA',
'South Carolina','SC','North Carolina','NC',
'South Dakota','SD','Iowa','IA',
'South Dakota','SD','Minnesota','MN',
'South Dakota','SD','Montana','MT',
'South Dakota','SD','Nebraska','NE',
'South Dakota','SD','North Dakota','ND',
'South Dakota','SD','Wyoming','WY',
'Tennessee','TN','Alabama','AL',
'Tennessee','TN','Arkansas','AR',
'Tennessee','TN','Georgia','GA',
'Tennessee','TN','Kentucky','KY',
'Tennessee','TN','Mississippi','MS',
'Tennessee','TN','Missouri','MO',
'Tennessee','TN','North Carolina','NC',
'Tennessee','TN','Virginia','VA',
'Texas','TX','Arkansas','AR',
'Texas','TX','Louisiana','LA',
'Texas','TX','New Mexico','NM',
'Texas','TX','Oklahoma','OK',
'Utah','UT','Arizona','AZ',
'Utah','UT','Colorado','CO',
'Utah','UT','Idaho','ID',
'Utah','UT','Nevada','NV',
'Utah','UT','New Mexico','NM',
'Utah','UT','Wyoming','WY',
'Vermont','VT','Massachusetts','MA',
'Vermont','VT','New Hampshire','NH',
'Vermont','VT','New York','NY',
'Virginia','VA','Kentucky','KY',
'Virginia','VA','Maryland','MD',
'Virginia','VA','North Carolina','NC',
'Virginia','VA','Tennessee','TN',
'Virginia','VA','West Virginia','WV',
'Washington','WA','Idaho','ID',
'Washington','WA','Oregon','OR',
'West Virginia','WV','Kentucky','KY',
'West Virginia','WV','Maryland','MD',
'West Virginia','WV','Ohio','OH',
'West Virginia','WV','Pennsylvania','PA',
'West Virginia','WV','Virginia','VA',
'Wisconsin','WI','Illinois','IL',
'Wisconsin','WI','Iowa','IA',
'Wisconsin','WI','Michigan','MI',
'Wisconsin','WI','Minnesota','MN',
'Wyoming','WY','Colorado','CO',
'Wyoming','WY','Idaho','ID',
'Wyoming','WY','Montana','MT',
'Wyoming','WY','Nebraska','NE',
'Wyoming','WY','South Dakota','SD',
'Wyoming','WY','Utah','UT'
)

neighbors_fullnames <- neighbors_raw %>%
  mutate(
    state = toupper(state),
    nbr   = toupper(nbr)
  ) %>%
  filter(nbr != "", !is.na(nbr)) %>%
  distinct(state, nbr) %>%
  group_by(state) %>%
  summarise(
    neighbour = paste(sort(unique(state_names[nbr])), collapse = ", "),
    .groups = "drop"
  ) %>%
  mutate(state = state_names[state]) %>%
  arrange(state)


neighbors_fullnames$state <- tolower(neighbors_fullnames$state)
neighbors_fullnames$neighbour <- tolower(neighbors_fullnames$neighbour)

neighbors_long <- neighbors_fullnames %>%
  mutate(
    state = tolower(state),
    neighbour = tolower(neighbour)
  ) %>%
  separate_rows(neighbour, sep = ",\\s*") %>%
  rename(neighbor_state = neighbour)
neighbors_long

```

### 4. Retrieve cattle dressing weights data from USDA NASS

```{r}

# Dressed carcass weights are converted to live weights using a 55% dressing percentage
dressing_pct <- 0.55

cattle_weight <- suppressMessages(
  readxl::read_xlsx(
    "/Users/sharaakat/Dropbox/akat_shara/DWFI_drought/GitHub/Data/Data/catsltr.xlsx",
    sheet = 2, skip = 3
  )
) %>%
  dplyr::select(1, 21) %>%                        # col 21 = cow dressed weight
  setNames(c("date", "dressed_wt")) %>%
  mutate(
    live_wt = dressed_wt / dressing_pct,
    year    = year(date),
    month   = month(date)
  ) %>%
  group_by(year) %>%
  summarize(mean_live_wt = mean(live_wt, na.rm = TRUE))
  
```

### 5. Retrieve weather data from PRISM

```{r}

# Download daily/monthly data from PRISM 
# num_cores <- detectCores()
# plan(multisession, workers = num_cores)
# #--- define a function to download and save PRISM data stacked by month ---#
# get_save_prism <- function(i, var_type) {
#
#   #++++++++++++++++++++++++++++++++++++
#   #+ Debug
#   #++++++++++++++++++++++++++++++++++++
#   # i <- 1
#   # var_type <- "tmax"
#
#   #++++++++++++++++++++++++++++++++++++
#   #+ Main
#   #++++++++++++++++++++++++++++++++++++
#   print(paste0("working on ", i))
#
#   temp_month <- month_year_data[i, month] # working month
#   temp_year <- month_year_data[i, year] # working year
#
#   options(prism.path = "/Volumes/Extreme SSD/UNL/DWFI/Drought/data/PRISM2022/")
#
#   #--- starting date of the working month-year ---#
#   start_date <- dmy(paste0("1/", temp_month, "/", temp_year))
#   #--- end date ---#
#   end_date <- start_date %m+% months(1) - 1
#
#   #--- download daily PRISM data for the working month-year ---#
#   get_prism_dailys(
#     type = var_type,
#     minDate = as.character(start_date),
#     maxDate = as.character(end_date),
#     keepZip = FALSE
#   )
#
#   #--- list of dates of the working month-year ---#
#   dates_ls <- seq(start_date, end_date, "days")
#
#   #--- remove dashes ---#
#   dates_prism_txt <- str_remove_all(dates_ls, "-")
#
#   #--- folder names ---#
#   folder_name <- paste0("PRISM_", var_type, "_stable_4kmD2_", dates_prism_txt, "_bil")
#   #--- the file name of the downloaded data ---#
#   file_name <- paste0("PRISM_", var_type, "_stable_4kmD2_", dates_prism_txt, "_bil.bil")
#   #--- complete path to the downloaded files ---#
#   file_path <- paste0("/Volumes/Extreme SSD/UNL/DWFI/Drought/data/PRISM2022/", folder_name, "/", file_name)
#
#   #--- combine all the PRISM files as a RasterStack ---#
#   temp_stars <-
#     stack(file_path) %>%
#     #--- convert to stars ---#
#     st_as_stars() %>%
#     #--- set the third dimension as data ---#
#     st_set_dimensions("band", values = dates_ls, name = "date")
#
#
#   #--- save the stars as an rds file ---#
#   saveRDS(
#     temp_stars,
#     paste0("/Volumes/Extreme SSD/UNL/DWFI/Drought/data/PRISM2022/PRISM_", var_type, "_y", temp_year, "_m", temp_month, ".rds")
#   )
#
# }
#
# month_year_data <- data.table::CJ(month = 1:12, year = c(2021, 2022))
#
#
# ## tmax, tmin, ppt, tmean
# future_lapply(
#   1:nrow(month_year_data),
#   function(x) get_save_prism(x, "tmin")
# )


# # Combine data from PRISM
# US_counties <- tigris::counties(state = "US", cb = TRUE) %>%
#   st_as_sf()
# 
# # 1981:1991; 1992:2002; 2003:2012; 2013:2020, 2021:2022
# month_year_data <- CJ(month = 1:12, year = 1981:2022)
# US_counties_data <- data.table()
# 
# temp_base <- 15  # base temperature for GDD calculation
# temp_up <- 31  # threshold temperature for EDD calculation
# 
# get_saved_PRISM_DT <- function(i, month_year_data, US_counties) {
#   # i = 1
#   temp_year <- month_year_data[i, year]
#   temp_month <- month_year_data[i, month]
# 
#   file_path_ppt <- sprintf("/Volumes/Extreme SSD/UNL/DWFI/Drought/data/PRISM_MONTHLY/PRISM_ppt_y%d_m%d.rds", temp_year, temp_month)
#   file_path_tmean <- sprintf("/Volumes/Extreme SSD/UNL/DWFI/Drought/data/PRISM_MONTHLY/PRISM_tmean_y%d_m%d.rds", temp_year, temp_month)
#   file_path_tmax <- sprintf("/Volumes/Extreme SSD/UNL/DWFI/Drought/data/PRISM_MONTHLY/PRISM_tmax_y%d_m%d.rds", temp_year, temp_month)
#   file_path_tmin <- sprintf("/Volumes/Extreme SSD/UNL/DWFI/Drought/data/PRISM_MONTHLY/PRISM_tmin_y%d_m%d.rds", temp_year, temp_month)
# 
#   cat("Year", temp_year, ", month", temp_month, "\n")
# 
#   saved_ppt <- readRDS(file_path_ppt)
#   saved_tmean <- readRDS(file_path_tmean)
#   saved_tmax <- readRDS(file_path_tmax)
#   saved_tmin <- readRDS(file_path_tmin)
# 
#   US_counties_sf <- st_transform(US_counties, st_crs(saved_ppt))
# 
#   a_ppt <- as.data.table(terra::extract(as(saved_ppt, "SpatRaster"), vect(US_counties_sf)))
#   a_tmean <- as.data.table(terra::extract(as(saved_tmean, "SpatRaster"), vect(US_counties_sf)))
#   a_tmax <- as.data.table(terra::extract(as(saved_tmax, "SpatRaster"), vect(US_counties_sf)))
#   a_tmin <- as.data.table(terra::extract(as(saved_tmin, "SpatRaster"), vect(US_counties_sf)))
# 
#   a_ppt_long <- melt(a_ppt, id.vars = "ID", variable.name = "Date", value.name = "ppt") %>%
#     .[, .(ppt = mean(ppt, na.rm = TRUE)), by = .(ID, Date)]
# 
#   # a_ppt_long <- melt(setDT(a_ppt), id.vars = "ID", variable.name = "Date", value.name = "ppt")[,
#   #      .(ppt = mean(ppt, na.rm = TRUE)), by = .(ID, Date)]
# 
#   a_tmean_long <- melt(a_tmean, id.vars = "ID", variable.name = "Date", value.name = "tmean") %>%
#     .[, .(tmean = mean(tmean, na.rm = TRUE)), by = .(ID, Date)]
#   a_tmax_long <- melt(a_tmax, id.vars = "ID", variable.name = "Date", value.name = "tmax") %>%
#     .[, .(tmax = mean(tmax, na.rm = TRUE)), by = .(ID, Date)]
#   a_tmin_long <- melt(a_tmin, id.vars = "ID", variable.name = "Date", value.name = "tmin") %>%
#     .[, .(tmin = mean(tmin, na.rm = TRUE)), by = .(ID, Date)]
# 
#   setkey(a_ppt_long, ID, Date)
#   setkey(a_tmean_long, ID, Date)
#   setkey(a_tmax_long, ID, Date)
#   setkey(a_tmin_long, ID, Date)
# 
#   # Joining the data.tableshe
#   a_long <- merge(a_ppt_long, a_tmean_long, by = c("ID", "Date"), all = TRUE)
#   a_long <- merge(a_long, a_tmax_long, by = c("ID", "Date"), all = TRUE)
#   a_long <- merge(a_long, a_tmin_long, by = c("ID", "Date"), all = TRUE)
# 
#   a_long[, Date := as.Date(paste0(gsub("date", "", Date), "01"), format = "%Y%m%d")]
#   a_long[, `:=`(month = month(Date), year = year(Date), day = day(Date))]
# 
#   US_counties_v_1 <- st_as_sf(vect(US_counties_sf)) %>%
#     mutate(ID = seq_len(nrow(.))) %>%
#     as.data.table()
# 
#   US_counties_v_1 <- merge(US_counties_v_1, a_long, by = "ID", all.x = TRUE)
#   US_counties_v_2 <- US_counties_v_1[, .(
#     tmin = mean(tmin, na.rm = TRUE),
#     tmax = mean(tmax, na.rm = TRUE),
#     tmean = mean(tmean, na.rm = TRUE),
#     ppt = sum(ppt, na.rm = TRUE)
#   ), by = .(year, month, NAME, STATE_NAME)]
# 
#   US_counties_v_2[, month_heat_index := (tmean / 5) ^ 1.514]
# 
#   US_counties_v_3 <- US_counties_v_2[, .(annual_heat_index = sum(month_heat_index, na.rm = TRUE)), by = .(year, NAME, STATE_NAME)]
#   US_counties_v_3[, a := 675 * 10^(-9) * annual_heat_index^3 - 771 * 10^(-7) * annual_heat_index^2 + 1792 * 10^(-5) * annual_heat_index + 0.49239]
# 
#   joined_weather_SPEI <- merge(US_counties_v_2, US_counties_v_3, by = c("year", "NAME", "STATE_NAME"), all.x = TRUE)
#   joined_weather_SPEI[, PET := ifelse(tmean <= 0, 0, 16 * ((10 * tmean) / annual_heat_index) ^ a)]
#   joined_weather_SPEI[, D := ppt - PET]
# 
# 
#   # Estimate GDD and EDD
#   US_counties_v_1[, Tavg := (tmax + tmin) / 2]
#   US_counties_v_1[, GDD := fifelse(Tavg < temp_base, 0,
#                                        fifelse(Tavg > temp_up, temp_up - temp_base,
#                                                Tavg - temp_base))]
# 
#   US_counties_v_1[, EDD := fifelse(tmax > temp_up, tmax - temp_up, 0)]
# 
#   US_counties_v_EDD_GDD <- US_counties_v_1[, .(
#     monthly_GDD = sum(GDD, na.rm = TRUE),
#     monthly_EDD = sum(EDD, na.rm = TRUE)
#   ), by = .(year, month, NAME, STATE_NAME)]
# 
# 
#   joined_weather <- merge(joined_weather_SPEI, US_counties_v_EDD_GDD,  by = c("year", "NAME", "STATE_NAME", "month"))
#   joined_weather <- joined_weather[, .(
#     avg_tmin = tmin,
#     avg_tmax = tmax,
#     avg_tmean = tmean,
#     sum_ppt = ppt,
#     monthly_GDD = monthly_GDD,
#     monthly_EDD = monthly_EDD,
#     monthly_D = sum(D, na.rm = TRUE)
#   ), by = .(year, month, NAME, STATE_NAME)]
# 
#   return(joined_weather)
# }
# 
# merged_data <- data.frame()
# merged_data <- lapply(
#   1:nrow(month_year_data),
#   function(x) get_saved_PRISM_DT(x, month_year_data, US_counties)
# )
# 
# combined <- do.call(rbind, merged_data)
# 
# saveRDS(
#   combined,
#   paste0("/Volumes/Extreme SSD/UNL/DWFI/Drought/data/Combined_NEW_GDD_EDD/", "15_combined_1981:2022.rds")
# )



# Region lookup
state_to_region <- function(state_name) {
  state_name <- tolower(state_name)
  dplyr::case_when(
    state_name %in% c("michigan", "wisconsin", "minnesota")                                                                              ~ "Lake States",
    state_name %in% c("iowa", "illinois", "indiana", "ohio", "missouri")                                                                 ~ "Corn Belt",
    state_name %in% c("north dakota", "south dakota", "nebraska", "kansas")                                                              ~ "Northern Plains",
    state_name %in% c("oklahoma", "texas")                                                                                               ~ "Southern Plains",
    state_name %in% c("arkansas", "louisiana", "mississippi")                                                                            ~ "Delta States",
    state_name %in% c("alabama", "south carolina", "georgia", "florida")                                                                 ~ "Southeast",
    state_name %in% c("west virginia", "kentucky", "virginia", "tennessee", "north carolina")                                            ~ "Appalachia",
    state_name %in% c("maine", "new hampshire", "vermont", "massachusetts", "rhode island",
                       "connecticut", "new york", "new jersey", "pennsylvania", "delaware", "maryland") ~ "Northeast",
    state_name %in% c("montana", "idaho", "wyoming", "colorado", "utah", "nevada", "arizona", "new mexico")                             ~ "Mountain",
    state_name %in% c("washington", "oregon", "california")                                                                              ~ "Pacific",
    TRUE ~ NA_character_
  )
}

# Region-specific grazing season start and end months (see Table 1 in manuscript)
region_grazing_months <- tibble::tribble(
  ~region,            ~start_month, ~end_month,
  "Appalachia",                  3,         10,
  "Corn Belt",                   4,         10,
  "Delta States",                4,         10,
  "Lake States",                 5,         10,
  "Mountain",                    5,         10,
  "Northeast",                   5,         10,
  "Northern Plains",             4,         10,
  "Pacific",                     3,         10,
  "Southeast",                   2,         10,
  "Southern Plains",             4,         10
)

# Helper: load pre-processed PRISM data, filter to grazing season, aggregate to annual
process_drought_data <- function(filepath) {
  readRDS(filepath) %>%
    mutate(
      monthly_D  = -monthly_D,                      # sign correction: D = PET - P
      STATE_NAME = tolower(STATE_NAME),
      NAME       = tolower(NAME),
      region     = state_to_region(STATE_NAME)
    ) %>%
    left_join(region_grazing_months, by = "region") %>%
    dplyr::filter(month >= start_month & month <= end_month) %>%  # grazing season only
    group_by(year, STATE_NAME, NAME) %>%
    summarize(
      total_D   = sum(monthly_D,   na.rm = TRUE),
      total_GDD = sum(monthly_GDD, na.rm = TRUE),
      total_EDD = sum(monthly_EDD, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    rename(state = STATE_NAME, county = NAME) %>%
    dplyr::select(year, state, county, total_D, total_GDD, total_EDD)
}

# Process and combine both periods
drought_final <- bind_rows(
  process_drought_data("/Users/sharaakat/Dropbox/akat_shara/DWFI_drought/GitHub/RobustnessCheck/Data/combined_1981_2020_NEW.rds"),
  process_drought_data("/Users/sharaakat/Dropbox/akat_shara/DWFI_drought/GitHub/RobustnessCheck/Data/combined_2021_2022_NEW.rds")
)


```


### 6. Retrieve Climate projections data 

```{r}

# 1. Functino to estimate weather variables:
US_counties_sf <- tigris::counties(cb = TRUE) %>%
  st_as_sf() %>%
  st_transform("EPSG:4326")

process_cmip6_data <- function(tmax_file, tmin_file, pr_file, scenario, temp_base = 12, temp_up = 31) {

  # Rotate to ensure correct alignment
  pr_raster <- terra::rotate(pr_file)
  tmin_raster <- terra::rotate(tmin_file)
  tmax_raster <- terra::rotate(tmax_file)
  
  # Set CRS to WGS84 explicitly
  crs(pr_raster) <- "EPSG:4326"
  crs(tmin_raster) <- "EPSG:4326"
  crs(tmax_raster) <- "EPSG:4326"
  
  # Assign time-based names to layers
  names(pr_raster) <- time(pr_raster)
  names(tmin_raster) <- time(tmin_raster)
  names(tmax_raster) <- time(tmax_raster)
  
  # Extract data
  pr_data <- as.data.table(terra::extract(pr_raster, vect(US_counties_sf)))
  tmin_data <- as.data.table(terra::extract(tmin_raster, vect(US_counties_sf)))
  tmax_data <- as.data.table(terra::extract(tmax_raster, vect(US_counties_sf)))
  
  # Reshape and aggregate data
  pr_data_ID <- data.table::melt(as.data.table(pr_data),
                                id.vars = "ID",
                                variable.name = "Date",
                                value.name = "ppt"
                                )[, .(ppt = mean(ppt, na.rm = TRUE)), by = .(ID, Date)]
  pr_data_ID[, Date_parsed := as.Date(as.character(Date))]
  pr_data_ID[, days_in_month := lubridate::days_in_month(Date_parsed)]
  pr_data_ID[, ppt := ppt * 86400 * days_in_month]
  pr_data_ID[, c("Date_parsed", "days_in_month") := NULL]
  
  
  tmin_data_ID <- data.table::melt(as.data.table(tmin_data),
                                  id.vars = "ID",
                                  variable.name = "Date",
                                  value.name = "tmin"
                                  )[, .(tmin = mean(tmin, na.rm = TRUE)), by = .(ID, Date)]

  tmax_data_ID <- data.table::melt(as.data.table(tmax_data),
                                    id.vars = "ID",
                                    variable.name = "Date",
                                    value.name = "tmax"
                                    )[, .(tmax = mean(tmax, na.rm = TRUE)), by = .(ID, Date)]
  
  setkey(pr_data_ID,   ID, Date)
  setkey(tmin_data_ID, ID, Date)
  setkey(tmax_data_ID, ID, Date)
  
  weather_data <- merge(pr_data_ID, tmin_data_ID, by = c("ID", "Date"))
  weather_data <- merge(weather_data, tmax_data_ID, by = c("ID", "Date"))
  
  cat("\nDone Phase 1")
  
  # Add year and month columns
  weather_data[, year := year(Date)]
  weather_data[, month := month(Date)]
  
  US_counties_v_1 <- st_as_sf(vect(US_counties_sf)) %>%
    mutate(ID = seq_len(nrow(.))) %>%
    as.data.table()
  
  US_counties_v_1 <- merge(US_counties_v_1, weather_data, by = "ID", all.x = TRUE)
  
  US_counties_v_1[, tmax := tmax - 273.15]
  US_counties_v_1[, tmin := tmin - 273.15]
  
  US_counties_v_1[, Tavg := (tmax + tmin) / 2]
    
  US_counties_v_1[, monthly_heat_index := fifelse(Tavg <= 0, 0, (Tavg / 5) ^ 1.514)]
  
  US_counties_annual <- US_counties_v_1[, .(
    annual_heat_index = sum(monthly_heat_index, na.rm = TRUE)
  ), by = .(year, NAME, STATE_NAME)]
  
  cat("\nDone Phase 2")
   
  US_counties_v_2 <-  merge(US_counties_v_1, US_counties_annual, by = c("year", "NAME", "STATE_NAME"), all.x = TRUE)
  
  US_counties_v_2[, a := 675 * 10^(-9) * annual_heat_index^3 - 771 * 10^(-7) * annual_heat_index^2 + 1792 * 10^(-5) * annual_heat_index + 0.49239]
  
  # joined_weather_SPEI <- merge(US_counties_v_1, US_counties_v_2, by = c("year", "NAME", "STATE_NAME"), all.x = TRUE)
  US_counties_v_2[, PET := ifelse(Tavg <= 0, 0, 16 * ((10 * Tavg) / annual_heat_index) ^ a)]
  US_counties_v_2[, D := PET - ppt]
  
  # joined_weather_SPEI[, Tavg := (tmax + tmin) / 2]
  US_counties_v_2[, GDD := fifelse(Tavg < temp_base, 0,
                                   fifelse(Tavg > temp_up, temp_up - temp_base,
                                           Tavg - temp_base))]
  
  US_counties_v_2[, EDD := fifelse(tmax > temp_up, tmax - temp_up, 0)]
  
   # if (scenario == "8.5") {
   #      US_counties_v_2[, GDD := GDD - 0.5 * EDD]  } # Adjust GDD by subtracting EDD for RCP 8.5
    joined_weather <- US_counties_v_2[, .(
    monthly_D = sum(D, na.rm = TRUE),
    monthly_GDD = sum(GDD, na.rm = TRUE),
    monthly_EDD = sum(EDD, na.rm = TRUE),
    monthly_ppt = sum(ppt, na.rm = TRUE),
    monthly_tmin = mean(tmin, na.rm = TRUE),
    monthly_tmax = mean(tmax, na.rm = TRUE)
  ), by = .(year, month, NAME, STATE_NAME)]
    
  cat("\nDone Phase 3")
      
  return(joined_weather)
}


# 2. Function to run through all realizations
process_cmip5_realization <- function(realization, base_dir, output_dir, temp_base = 12, temp_up = 31) {
  
  # Construct file paths dynamically based on realization number
  tmax_4.5 <- rast(file.path(base_dir, sprintf("r%02d/tasmax_Amon_CanESM5_rcp45-cmip5_r%di1p1f1_gn_200601-210012.nc", realization, realization)))
  tmin_4.5 <- rast(file.path(base_dir, sprintf("r%02d/tasmin_Amon_CanESM5_rcp45-cmip5_r%di1p1f1_gn_200601-210012.nc", realization, realization)))
  pr_4.5   <- rast(file.path(base_dir, sprintf("r%02d/pr_Amon_CanESM5_rcp45-cmip5_r%di1p1f1_gn_200601-210012.nc", realization, realization)))

  tmax_8.5 <- rast(file.path(base_dir, sprintf("r%02d/tasmax_Amon_CanESM5_rcp85-cmip5_r%di1p1f1_gn_200601-210012.nc", realization, realization)))
  tmin_8.5 <- rast(file.path(base_dir, sprintf("r%02d/tasmin_Amon_CanESM5_rcp85-cmip5_r%di1p1f1_gn_200601-210012.nc", realization, realization)))
  pr_8.5   <- rast(file.path(base_dir, sprintf("r%02d/pr_Amon_CanESM5_rcp85-cmip5_r%di1p1f1_gn_200601-210012.nc", realization, realization)))

  # Process CMIP5 data for each scenario
  file_45 <- process_cmip6_data(tmax_4.5, tmin_4.5, pr_4.5, "4.5", temp_base, temp_up)
  file_85 <- process_cmip6_data(tmax_8.5, tmin_8.5, pr_8.5, "8.5", temp_base, temp_up)


  # Save results with realization number
  saveRDS(file_45, file.path(output_dir, sprintf("Weather_r%02d_4.5.rds", realization)))
  cat("\nFinished phase 4")
  saveRDS(file_85, file.path(output_dir, sprintf("Weather_r%02d_8.5.rds", realization)))
  cat("\nFinished phase 5")

  cat(sprintf("Processed and saved results for realization r%02d\n", realization))
}

# Example Usage: Run for multiple realizations
base_dir <- "/Volumes/Extreme SSD/UNL/DWFI/Drought/data/Climate_Projection_2"
output_dir <- "/Volumes/Extreme SSD/UNL/DWFI/Drought/data/Climate_Projection_2"
for (r in 1:10) {
  folder <- file.path(base_dir, sprintf("r%02d", r))
  files  <- list.files(folder, pattern = "tasmax")
  cat("Realization", r, ":", files, "\n")
}


# Run the function for realizations 1 through 10
for (realization in 1:10) {
  process_cmip5_realization(realization, base_dir, output_dir)
}

# data_2 <- readRDS("/Volumes/Extreme SSD/UNL/DWFI/Drought/data/Climate_Projection_2/Weather_r01_4.5.rds")
# data_3 <- readRDS("/Volumes/Extreme SSD/UNL/DWFI/Drought/data/Climate_Projection_2/Weather_r01_8.5.rds")


```


### 7.Retrieve USDM data

```{r}

USDM_data_weekly <- read.csv("/Users/sharaakat/Dropbox/akat_shara/DWFI_drought/data/dm_export_20000101_20260126.csv")
USDM_data_weekly2 <- USDM_data_weekly %>%
  mutate(
    state  = state_names[State],          # if State is an abbreviation like "AK"
    region = state_to_region(state),
    date   = ymd(MapDate),
    month  = month(date),
    year   = year(date)
  ) %>%
  left_join(region_grazing_months, by = "region") %>% na.omit()


USDM_annual <- USDM_data_weekly2 %>%
  filter(month >= start_month & month <= end_month) %>%
  group_by(County, state, year) %>%
  summarise(
    D0_weeks = sum(D0/100, na.rm = TRUE),
    D1_weeks = sum(D1/100, na.rm = TRUE),
    D2_weeks = sum(D2/100, na.rm = TRUE),
    D3_weeks = sum(D3/100, na.rm = TRUE),
    D4_weeks = sum(D4/100, na.rm = TRUE),
    .groups = "drop"
  )

USDM_annual$county <- tolower(USDM_annual$County)
USDM_annual$state <- tolower(USDM_annual$state)
USDM_annual$county <- trimws(gsub("\\s+county\\b", "", tolower(USDM_annual$County)))


```


### 8.Retrieve PSDI data

```{r}

counties_us <- counties(cb = TRUE, year = 2022, class = "sf") %>%
  st_transform(4326) %>%
  filter(!STATEFP %in% c("02", "15", "72")) %>%  # remove AK, HI, Puerto Rico
  mutate(
    fips = GEOID,
    county_name = NAME,
    statefp = STATEFP
  ) %>%
  select(fips, statefp, county_name, geometry)


get_county_pdsi_year <- function(year) {
  message("Processing year: ", year)
  
  url <- paste0(
    "https://climate.northwestknowledge.net/TERRACLIMATE-DATA/",
    "TerraClimate_pdsi_", year, ".nc"
  )
  
  nc_file <- file.path(tempdir(), paste0("TerraClimate_pdsi_", year, ".nc"))
  
  if (!file.exists(nc_file)) {
    download.file(url, nc_file, mode = "wb")
  }
  
  r <- rast(nc_file)
  
  # Make sure raster layers have clear month names
  names(r) <- paste0("pdsi_", sprintf("%02d", 1:nlyr(r)))
  
  # Convert counties to same CRS as raster
  counties_raster_crs <- st_transform(counties_us, crs(r))
  
  # Crop raster to counties
  r_us <- crop(r, vect(counties_raster_crs))
  
  # Extract county-level monthly means
  extracted <- exact_extract(r_us, counties_raster_crs, "mean")
  
  # Force extracted columns to month names
  names(extracted) <- paste0("pdsi_", sprintf("%02d", 1:ncol(extracted)))
  
  county_monthly <- bind_cols(
    st_drop_geometry(counties_us),
    as_tibble(extracted)
  ) %>%
    pivot_longer(
      cols = starts_with("pdsi_"),
      names_to = "month",
      values_to = "pdsi"
    ) %>%
    mutate(
      year = year,
      month = as.integer(str_remove(month, "pdsi_"))
    )
  
  return(county_monthly)
}

pdsi_county_monthly <- map_dfr(1981:2022, get_county_pdsi_year)

pdsi_county_monthly_clean <- pdsi_county_monthly %>%
  mutate(
    statefp = substr(fips, 1, 2)
  ) %>%
  left_join(
    tigris::fips_codes %>%
      distinct(state_code, state_name),
    by = c("statefp" = "state_code")
  ) %>%
  rename(
    state = state_name,
    county = county_name,
    index = pdsi
  ) %>%
  select(state, county, year, month, index)

pdsi_county_monthly_clean$state <- tolower(pdsi_county_monthly_clean$state)
pdsi_county_monthly_clean$county <- tolower(pdsi_county_monthly_clean$county)
write_csv(pdsi_county_monthly_clean, "/Users/sharaakat/Dropbox/akat_shara/DWFI_drought/PDSI_1981_2022_monthly.csv")

```





