---
title: "Data Collection"
date: "`r Sys.Date()`"
format: html
execute:
  echo: true
---

### ============================================================

### Data Collection Script

### This script collects and processes raw data from various sources

### Sources include:

### - USDA Census of Agriculture: Pastureland and cattle inventory data

### - USDA NASS: Hay Production Data

### - USDA NASS: Cattle weights

### - PRISM: Weather Data

### - MACA dataset: Climate Projections

### ============================================================

```{r}
pacman::p_load(
  pacman,
  stars, # spatiotemporal data handling
  terra, # raster data handling
  raster, # raster data handling
  sf, # vector data handling
  dplyr, # data wrangling
  tidyverse, 
  fixest,
  parallel,
  foreach,
  doParallel,
  cubelyr,
  tigris,
  zoo,
  SCI,
  spei,
  purrr,
  datatable,
  ncdf4,
  stringr, # string manipulation
  lubridate, # dates handling
  data.table, # data wrangling
  tidyr, # reshape
  tidyUSDA, # download USDA NASS data
  keyring, # API key management
  FedData, # download Daymet data
  daymetr, # download Daymet data
  ggplot2, # make maps
  tmap, # make maps
  future.apply, # parallel processing
  CropScapeR, # download CDL data
  prism, # download PRISM data
  exactextractr # extract raster values to sf
)
rm(list = ls())

# Setting USDA-NASS key
usdarnass::nass_set_key(key = "79F68508-887F-3423-A3EE-F60AB7DFB3AE")
```

### 1. Retrieve USDA NASS pastureland data from USDA Census of Agriculture

```{r}
#| echo: false

## PASTURE LAND - from USDA NASS quickStats
pasture_land_total <- usdarnass::nass_data(source_desc = "CENSUS", sector_desc = "ECONOMICS", group_desc = "FARMS & LAND & ASSETS",
                                      commodity_desc = "AG LAND", statisticcat_desc = "AREA", short_desc = "AG LAND, PASTURELAND - ACRES",
                                      domain_desc = "TOTAL", agg_level_desc = "NATIONAL", year = "1990<=") %>% 
  dplyr::select(., state_name, county_name, year, Value) 

pasture_land_total <- rename(pasture_land_total, state = state_name)
pasture_land_total <- rename(pasture_land_total, county = county_name)
pasture_land_total <- rename(pasture_land_total, pastureland_all_types = Value)


# PASTURE LAND - from PDFs (before 1997)
pasture_data_PDF <- readxl::read_xlsx('/Users/sharaakat/Dropbox/akat_shara/DWFI_drought/GitHub/Data/Data/pasture_data_FINAL.xlsx')


# Join all PASTURELAND data
pasture_data_ALL <- rbind(pasture_data_PDF, pasture_land_total)
pasture_data_ALL$year <- as.numeric(as.character(pasture_data_ALL$year))
pasture_data_ALL$pastureland_all_types <- gsub(",", "", pasture_data_ALL$pastureland_all_types)
pasture_data_ALL$pastureland_all_types <- as.numeric(as.character(pasture_data_ALL$pastureland_all_types))
pasture_data_ALL$state <- tolower(pasture_data_ALL$state)
pasture_data_ALL$county <- tolower(pasture_data_ALL$county)


```

### 2. Retrieve USDA NASS cattle inventory data from USDA Census of Agriculture

```{r}

## CATTLE INV - from USDA NASS quickStats
beef_inv_total <- usdarnass::nass_data(source_desc = "CENSUS", sector_desc = "ANIMALS & PRODUCTS", group_desc = "LIVESTOCK",
                                      commodity_desc = "CATTLE", statisticcat_desc = "INVENTORY", short_desc = "CATTLE, COWS, BEEF - INVENTORY",
                                      domain_desc = "TOTAL",  agg_level_desc = "NATIONAL", year = "1990<=") %>% 
  dplyr::select(., state_name, county_name, year, Value) %>% 
  rename(beef_inv = Value,
         state = state_name,
         county = county_name)


# CATTLE INV - from PDFs (before 1997)
beef_inv_PDF <- readxl::read_xlsx('/Users/sharaakat/Dropbox/akat_shara/DWFI_drought/GitHub/Data/Data/county_beefinv.xlsx')


# Join all CATTLE INV. data
beef_inv <- rbind(beef_inv_PDF, beef_inv_total)
beef_inv$year <- as.numeric(as.character(beef_inv$year))
beef_inv$beef_inv <- gsub(",", "", beef_inv$beef_inv)
beef_inv$beef_inv <- as.numeric(as.character(beef_inv$beef_inv))
beef_inv$state <- tolower(beef_inv$state)
beef_inv$county <- tolower(beef_inv$county)

```

### 3. Retrieve USDA NASS hay production data

```{r}
## HAY PRODUCTION (lbs) DATA
state_names <- c(
  AL = "Alabama", AK = "Alaska", AZ = "Arizona", AR = "Arkansas", CA = "California", 
  CO = "Colorado", CT = "Connecticut", DE = "Delaware", FL = "Florida", GA = "Georgia", 
  HI = "Hawaii", ID = "Idaho", IL = "Illinois", IN = "Indiana", IA = "Iowa", 
  KS = "Kansas", KY = "Kentucky", LA = "Louisiana", ME = "Maine", MD = "Maryland", 
  MA = "Massachusetts", MI = "Michigan", MN = "Minnesota", MS = "Mississippi", 
  MO = "Missouri", MT = "Montana", NE = "Nebraska", NV = "Nevada", NH = "New Hampshire", 
  NJ = "New Jersey", NM = "New Mexico", NY = "New York", NC = "North Carolina", 
  ND = "North Dakota", OH = "Ohio", OK = "Oklahoma", OR = "Oregon", PA = "Pennsylvania", 
  RI = "Rhode Island", SC = "South Carolina", SD = "South Dakota", TN = "Tennessee", 
  TX = "Texas", UT = "Utah", VT = "Vermont", VA = "Virginia", WA = "Washington", 
  WV = "West Virginia", WI = "Wisconsin", WY = "Wyoming"
)

# All hat stocks, 1000 tones
hay_all_stocks_1000t <- readxl::read_xlsx("/Users/sharaakat/Dropbox/akat_shara/DWFI_drought/GitHub/RobustnessCheck/Data/StateHaySupplies.xlsx", sheet = 2, skip = 2)

hay_all_stocks_1000t_2 <- pivot_longer(hay_all_stocks_1000t, cols = -year, names_to = "state", values_to = "hay_stock_1000t") %>% 
  mutate(state = state_names[state])

hay_all_stocks_1000t_2$state <- tolower(hay_all_stocks_1000t_2$state)

hay_all_stocks_1000t_3 <- hay_all_stocks_1000t_2 %>%
  group_by(state) %>%
  mutate(hay_stock_1000t = dplyr::lead(hay_stock_1000t, n = 1)) %>%
  ungroup() 

hay_all_stocks_1000t_national <- hay_all_stocks_1000t_3 %>% 
  group_by(year) %>% 
  summarize(hay_stock_1000t = mean(hay_stock_1000t, na.rm = TRUE))
```

### 4. Retrieve cattle dressing weights data from USDA NASS

```{r}
cattle_weight <- readxl::read_xlsx("Data/Data/catsltr.xlsx", sheet = 2, skip = 3) %>% 
  dplyr::select(1, 18) %>% 
  `colnames<-`(c("date", "dressed_wt")) %>% 
  dplyr::mutate(live_wt = dressed_wt/dressing_pct,
                year = year(date), month = month(date)) %>% 
  group_by(year) %>% 
  summarize(mean_live_wt = mean(live_wt, na.rm = TRUE))
```

### 5. Retrieve weather data from PRISM

```{r}

# US_counties <- tigris::counties(state = "US", cb = TRUE) %>%
#   st_as_sf()
# 
# # 1981:1991; 1992:2002; 2003:2012; 2013:2020
# month_year_data <- CJ(month = 1:12, year = 2006)
# US_counties_data <- data.table()
# 
# temp_base <- 12  # base temperature for GDD calculation
# temp_up <- 31  # threshold temperature for EDD calculation
# 
# get_saved_PRISM_DT <- function(i, month_year_data, US_counties) {
#   # i = 1
#   temp_year <- month_year_data[i, year]
#   temp_month <- month_year_data[i, month]
# 
#   file_path_ppt <- sprintf("/Volumes/Extreme SSD/UNL/DWFI/data/PRISM_MONTHLY/PRISM_ppt_y%d_m%d.rds", temp_year, temp_month)
#   file_path_tmean <- sprintf("/Volumes/Extreme SSD/UNL/DWFI/data/PRISM_MONTHLY/PRISM_tmean_y%d_m%d.rds", temp_year, temp_month)
#   file_path_tmax <- sprintf("/Volumes/Extreme SSD/UNL/DWFI/data/PRISM_MONTHLY/PRISM_tmax_y%d_m%d.rds", temp_year, temp_month)
#   file_path_tmin <- sprintf("/Volumes/Extreme SSD/UNL/DWFI/data/PRISM_MONTHLY/PRISM_tmin_y%d_m%d.rds", temp_year, temp_month)
# 
#   cat("Year", temp_year, ", month", temp_month, "\n")
# 
#   saved_ppt <- readRDS(file_path_ppt)
#   saved_tmean <- readRDS(file_path_tmean)
#   saved_tmax <- readRDS(file_path_tmax)
#   saved_tmin <- readRDS(file_path_tmin)
# 
#   US_counties_sf <- st_transform(US_counties, st_crs(saved_ppt))
# 
#   a_ppt <- as.data.table(terra::extract(as(saved_ppt, "SpatRaster"), vect(US_counties_sf)))
#   a_tmean <- as.data.table(terra::extract(as(saved_tmean, "SpatRaster"), vect(US_counties_sf)))
#   a_tmax <- as.data.table(terra::extract(as(saved_tmax, "SpatRaster"), vect(US_counties_sf)))
#   a_tmin <- as.data.table(terra::extract(as(saved_tmin, "SpatRaster"), vect(US_counties_sf)))
# 
#   a_ppt_long <- melt(a_ppt, id.vars = "ID", variable.name = "Date", value.name = "ppt") %>%
#     .[, .(ppt = mean(ppt, na.rm = TRUE)), by = .(ID, Date)]
# 
#   # a_ppt_long <- melt(setDT(a_ppt), id.vars = "ID", variable.name = "Date", value.name = "ppt")[,
#   #      .(ppt = mean(ppt, na.rm = TRUE)), by = .(ID, Date)]
# 
#   a_tmean_long <- melt(a_tmean, id.vars = "ID", variable.name = "Date", value.name = "tmean") %>%
#     .[, .(tmean = mean(tmean, na.rm = TRUE)), by = .(ID, Date)]
#   a_tmax_long <- melt(a_tmax, id.vars = "ID", variable.name = "Date", value.name = "tmax") %>%
#     .[, .(tmax = mean(tmax, na.rm = TRUE)), by = .(ID, Date)]
#   a_tmin_long <- melt(a_tmin, id.vars = "ID", variable.name = "Date", value.name = "tmin") %>%
#     .[, .(tmin = mean(tmin, na.rm = TRUE)), by = .(ID, Date)]
# 
#   setkey(a_ppt_long, ID, Date)
#   setkey(a_tmean_long, ID, Date)
#   setkey(a_tmax_long, ID, Date)
#   setkey(a_tmin_long, ID, Date)
# 
#   # Joining the data.tables
#   a_long <- merge(a_ppt_long, a_tmean_long, by = c("ID", "Date"), all = TRUE)
#   a_long <- merge(a_long, a_tmax_long, by = c("ID", "Date"), all = TRUE)
#   a_long <- merge(a_long, a_tmin_long, by = c("ID", "Date"), all = TRUE)
# 
#   a_long[, Date := as.Date(gsub("date", "", Date))]
#   a_long[, `:=`(month = month(Date), year = year(Date))]
# 
#   US_counties_v_1 <- st_as_sf(vect(US_counties_sf)) %>%
#     mutate(ID = seq_len(nrow(.))) %>%
#     as.data.table()
# 
#   US_counties_v_1 <- merge(US_counties_v_1, a_long, by = "ID", all.x = TRUE)
#   US_counties_v_1 <- US_counties_v_1[, .(
#     tmin = mean(tmin, na.rm = TRUE),
#     tmax = mean(tmax, na.rm = TRUE),
#     tmean = mean(tmean, na.rm = TRUE),
#     ppt = sum(ppt, na.rm = TRUE)
#   ), by = .(year, month, NAME, STATE_NAME)]
# 
# 
#   US_counties_v_1[, month_heat_index := (tmean / 5) ^ 1.514]
# 
#   US_counties_v_2 <- US_counties_v_1[, .(annual_heat_index = sum(month_heat_index, na.rm = TRUE)), by = .(year, NAME, STATE_NAME)]
#   US_counties_v_2[, a := 675 * 10^(-9) * annual_heat_index^3 - 771 * 10^(-7) * annual_heat_index^2 + 1792 * 10^(-5) * annual_heat_index + 0.49239]
# 
#   joined_weather_SPEI <- merge(US_counties_v_1, US_counties_v_2, by = c("year", "NAME", "STATE_NAME"), all.x = TRUE)
#   joined_weather_SPEI[, PET := ifelse(tmean <= 0, 0, 16 * ((10 * tmean) / annual_heat_index) ^ a)]
#   joined_weather_SPEI[, D := ppt - PET]
# 
#   joined_weather_SPEI[, Tavg := (tmax + tmin) / 2]
#   joined_weather_SPEI[, GDD := fifelse(Tavg < temp_base, 0,
#                                        fifelse(Tavg > temp_up, temp_up - temp_base,
#                                                Tavg - temp_base))]
# 
#   joined_weather_SPEI[, EDD := fifelse(tmax > temp_up, tmax - temp_up, 0)]
# 
#   joined_weather <- joined_weather_SPEI[, .(
#     avg_tmin = tmin,
#     avg_tmax = tmax,
#     avg_tmean = tmean,
#     sum_ppt = ppt,
#     monthly_D = sum(D, na.rm = TRUE),
#     monthly_GDD = sum(GDD, na.rm = TRUE),
#     monthly_EDD = sum(EDD, na.rm = TRUE)
#   ), by = .(year, month, NAME, STATE_NAME)]
# 
#   return(joined_weather)
# }
# 
# merged_data_1981_2020 <- data.frame()
# merged_data_1981_2020 <- lapply(
#   1:nrow(month_year_data),
#   function(x) get_saved_PRISM_DT(x, month_year_data, US_counties)
# )
# 
# combined_1981_2020 <- do.call(rbind, merged_data_1981_2020)
# 
# saveRDS(
#   combined_1981_2020,
#   paste0("/Volumes/Extreme SSD/UNL/DWFI /data/PRISM_MONTHLY_COMBINED_10_31/", "combined_1981_2020.rds")
# )

drought_final <- readRDS("/Users/sharaakat/Dropbox/akat_shara/DWFI_drought/GitHub/RobustnessCheck/Data/combined_1981_2020_NEW.rds")

growing_season_data <- drought_final %>%
  filter(month >= 3 & month <= 10)  # Select only March to October

# Aggregate to yearly level
yearly_drought_data <- growing_season_data %>%
  group_by(year, STATE_NAME, NAME) %>%  # Group by year, state, and county
  summarize(
    total_D = sum(D, na.rm = TRUE),  # Total drought index
    total_GDD = sum(GDD, na.rm = TRUE),  # Total Growing Degree Days
    total_EDD = sum(EDD, na.rm = TRUE)  # Total Extreme Degree Days
  ) %>% 
  rename(state = STATE_NAME,
         county = NAME) %>% 
  dplyr::select(year, state, county, total_D, total_GDD, total_EDD)

yearly_drought_data$state <- tolower(yearly_drought_data$state)
yearly_drought_data$county <- tolower(yearly_drought_data$county)

```

### 6. Retrieve Climate projections data from MACA

```{r}
# 1. First we need to download the MACA projected climate data for several different Global Climate Models (GCMs)

# ## 1. bcc-csm1-1 (China)
# ## 1.1. pr DATA
# # 4.5, pr: 2006-2099, monthly
# urls_pr_45_1 <- c(
#   "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1/macav2livneh_pr_bcc-csm1-1_r1i1p1_rcp45_2006_2025_CONUS_monthly.nc",
#   "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1/macav2livneh_pr_bcc-csm1-1_r1i1p1_rcp45_2026_2045_CONUS_monthly.nc",
#   "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1/macav2livneh_pr_bcc-csm1-1_r1i1p1_rcp45_2046_2065_CONUS_monthly.nc",
#   "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1/macav2livneh_pr_bcc-csm1-1_r1i1p1_rcp45_2066_2085_CONUS_monthly.nc",
#   "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1/macav2livneh_pr_bcc-csm1-1_r1i1p1_rcp45_2086_2099_CONUS_monthly.nc"
# )
# 
# # 8.5, pr: 2006-2099, monthly
# urls_pr_85_1 <- c(
#   "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1/macav2livneh_pr_bcc-csm1-1_r1i1p1_rcp85_2006_2025_CONUS_monthly.nc",
#   "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1/macav2livneh_pr_bcc-csm1-1_r1i1p1_rcp85_2026_2045_CONUS_monthly.nc",
#   "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1/macav2livneh_pr_bcc-csm1-1_r1i1p1_rcp85_2046_2065_CONUS_monthly.nc",
#   "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1/macav2livneh_pr_bcc-csm1-1_r1i1p1_rcp85_2066_2085_CONUS_monthly.nc",
#   "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1/macav2livneh_pr_bcc-csm1-1_r1i1p1_rcp85_2086_2099_CONUS_monthly.nc"
# )


## 2. bcc-csm1-1-m (China)
## 1.1. pr DATA
# 4.5, pr: 2006-2099, monthly
urls_pr_45_1 <- c(
  "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1-m/macav2livneh_pr_bcc-csm1-1-m_r1i1p1_rcp45_2006_2025_CONUS_monthly.nc",
  "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1-m/macav2livneh_pr_bcc-csm1-1-m_r1i1p1_rcp45_2026_2045_CONUS_monthly.nc",
  "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1-m/macav2livneh_pr_bcc-csm1-1-m_r1i1p1_rcp45_2046_2065_CONUS_monthly.nc",
  "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1-m/macav2livneh_pr_bcc-csm1-1-m_r1i1p1_rcp45_2066_2085_CONUS_monthly.nc",
  "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1-m/macav2livneh_pr_bcc-csm1-1-m_r1i1p1_rcp45_2086_2099_CONUS_monthly.nc"
)

# 8.5, pr: 2006-2099, monthly
urls_pr_85_1 <- c(
  "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1-m/macav2livneh_pr_bcc-csm1-1-m_r1i1p1_rcp85_2006_2025_CONUS_monthly.nc",
  "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1-m/macav2livneh_pr_bcc-csm1-1-m_r1i1p1_rcp85_2026_2045_CONUS_monthly.nc",
  "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1-m/macav2livneh_pr_bcc-csm1-1-m_r1i1p1_rcp85_2046_2065_CONUS_monthly.nc",
  "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1-m/macav2livneh_pr_bcc-csm1-1-m_r1i1p1_rcp85_2066_2085_CONUS_monthly.nc",
  "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1-m/macav2livneh_pr_bcc-csm1-1-m_r1i1p1_rcp85_2086_2099_CONUS_monthly.nc"
)



urls_pr_45_1 <- c(
  "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1-m/macav2livneh_pr_bcc-csm1-1-m_r1i1p1_rcp45_2006_2025_CONUS_monthly.nc",
  "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1-m/macav2livneh_pr_bcc-csm1-1-m_r1i1p1_rcp45_2026_2045_CONUS_monthly.nc",
  "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1-m/macav2livneh_pr_bcc-csm1-1-m_r1i1p1_rcp45_2046_2065_CONUS_monthly.nc",
  "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1-m/macav2livneh_pr_bcc-csm1-1-m_r1i1p1_rcp45_2066_2085_CONUS_monthly.nc",
  "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1-m/macav2livneh_pr_bcc-csm1-1-m_r1i1p1_rcp45_2086_2099_CONUS_monthly.nc"
)

# 8.5, pr: 2006-2099, monthly
urls_pr_85_1 <- c(
  "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1-m/macav2livneh_pr_bcc-csm1-1-m_r1i1p1_rcp85_2006_2025_CONUS_monthly.nc",
  "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1-m/macav2livneh_pr_bcc-csm1-1-m_r1i1p1_rcp85_2026_2045_CONUS_monthly.nc",
  "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1-m/macav2livneh_pr_bcc-csm1-1-m_r1i1p1_rcp85_2046_2065_CONUS_monthly.nc",
  "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1-m/macav2livneh_pr_bcc-csm1-1-m_r1i1p1_rcp85_2066_2085_CONUS_monthly.nc",
  "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1-m/macav2livneh_pr_bcc-csm1-1-m_r1i1p1_rcp85_2086_2099_CONUS_monthly.nc"
)

# Loop to download and extract each file
for (url in urls_pr_45_1) {
  # Extract year range from the URL for naming the file
  year_range <- gsub(".*_([0-9]{4}_[0-9]{4})_.*", "\\1", url)
  
  # Set destination file name
  destfile <- paste0("/Volumes/Extreme SSD/UNL/DWFI/projection/data/2/pr4.5_", year_range, ".nc")
  
  # Increase timeout in case of large files
  options(timeout = max(600, getOption("timeout")))
  
  # Download the file
  download.file(url, destfile = destfile, method = "auto", mode = "wb")
  
  # Load the .nc file into a raster object
  assign(paste0("pr4.5", year_range), terra::rast(destfile))
  
  # Print confirmation of download and extraction
  print(paste("Downloaded and loaded:", destfile))
}

# Loop to download and extract each file
for (url in urls_pr_85_1) {
  # Extract year range from the URL for naming the file
  year_range <- gsub(".*_([0-9]{4}_[0-9]{4})_.*", "\\1", url)
  
  # Set destination file name
  destfile <- paste0("/Volumes/Extreme SSD/UNL/DWFI/projection/data/2/pr8.5_", year_range, ".nc")
  
  # Increase timeout in case of large files
  options(timeout = max(600, getOption("timeout")))
  
  # Download the file
  download.file(url, destfile = destfile, method = "auto", mode = "wb")
  
  # Load the .nc file into a raster object
  assign(paste0("pr8.5_", year_range), terra::rast(destfile))
  
  # Print confirmation of download and extraction
  print(paste("Downloaded and loaded:", destfile))
}



## 1.2. tmin DATA
# 4.5, tmin: 2006-2099, monthly
# urls_tmin_45_1 <- c(
#   "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1/macav2livneh_tasmin_bcc-csm1-1_r1i1p1_rcp45_2006_2025_CONUS_monthly.nc",
#   "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1/macav2livneh_tasmin_bcc-csm1-1_r1i1p1_rcp45_2026_2045_CONUS_monthly.nc",
#   "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1/macav2livneh_tasmin_bcc-csm1-1_r1i1p1_rcp45_2046_2065_CONUS_monthly.nc",
#   "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1/macav2livneh_tasmin_bcc-csm1-1_r1i1p1_rcp45_2066_2085_CONUS_monthly.nc",
#   "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1/macav2livneh_tasmin_bcc-csm1-1_r1i1p1_rcp45_2086_2099_CONUS_monthly.nc"
# )
# 
# 
# # 8.5, tmin: 2006-2099, monthly
# urls_tmin_85_1 <- c(
#   "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1/macav2livneh_tasmin_bcc-csm1-1_r1i1p1_rcp85_2006_2025_CONUS_monthly.nc",
#   "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1/macav2livneh_tasmin_bcc-csm1-1_r1i1p1_rcp85_2026_2045_CONUS_monthly.nc",
#   "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1/macav2livneh_tasmin_bcc-csm1-1_r1i1p1_rcp85_2046_2065_CONUS_monthly.nc",
#   "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1/macav2livneh_tasmin_bcc-csm1-1_r1i1p1_rcp85_2066_2085_CONUS_monthly.nc",
#   "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1/macav2livneh_tasmin_bcc-csm1-1_r1i1p1_rcp85_2086_2099_CONUS_monthly.nc"
# )

urls_tmin_45_1 <- c(
  "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1-m/macav2livneh_tasmin_bcc-csm1-1-m_r1i1p1_rcp45_2006_2025_CONUS_monthly.nc",
  "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1-m/macav2livneh_tasmin_bcc-csm1-1-m_r1i1p1_rcp45_2026_2045_CONUS_monthly.nc",
  "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1-m/macav2livneh_tasmin_bcc-csm1-1-m_r1i1p1_rcp45_2046_2065_CONUS_monthly.nc",
  "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1-m/macav2livneh_tasmin_bcc-csm1-1-m_r1i1p1_rcp45_2066_2085_CONUS_monthly.nc",
  "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1-m/macav2livneh_tasmin_bcc-csm1-1-m_r1i1p1_rcp45_2086_2099_CONUS_monthly.nc"
)


# 8.5, tmin: 2006-2099, monthly
urls_tmin_85_1 <- c(
  "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1-m/macav2livneh_tasmin_bcc-csm1-1-m_r1i1p1_rcp85_2006_2025_CONUS_monthly.nc",
  "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1-m/macav2livneh_tasmin_bcc-csm1-1-m_r1i1p1_rcp85_2026_2045_CONUS_monthly.nc",
  "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1-m/macav2livneh_tasmin_bcc-csm1-1-m_r1i1p1_rcp85_2046_2065_CONUS_monthly.nc",
  "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1-m/macav2livneh_tasmin_bcc-csm1-1-m_r1i1p1_rcp85_2066_2085_CONUS_monthly.nc",
  "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1-m/macav2livneh_tasmin_bcc-csm1-1-m_r1i1p1_rcp85_2086_2099_CONUS_monthly.nc"
)

# Loop to download and extract each file
for (url in urls_tmin_45_1) {
  # Extract year range from the URL for naming the file
  year_range <- gsub(".*_([0-9]{4}_[0-9]{4})_.*", "\\1", url)
  
  # Set destination file name
  destfile <- paste0("/Volumes/Extreme SSD/UNL/DWFI/projection/data/2/tmin4.5_", year_range, ".nc")
  
  # Increase timeout in case of large files
  options(timeout = max(600, getOption("timeout")))
  
  # Download the file
  download.file(url, destfile = destfile, method = "auto", mode = "wb")
  
  # Load the .nc file into a raster object
  assign(paste0("tmin4.5_", year_range), terra::rast(destfile))
  
  # Print confirmation of download and extraction
  print(paste("Downloaded and loaded:", destfile))
}


# Loop to download and extract each file
for (url in urls_tmin_85_1) {
  # Extract year range from the URL for naming the file
  year_range <- gsub(".*_([0-9]{4}_[0-9]{4})_.*", "\\1", url)
  
  # Set destination file name
  destfile <- paste0("/Volumes/Extreme SSD/UNL/DWFI/projection/data/2/tmin8.5_", year_range, ".nc")
  
  # Increase timeout in case of large files
  options(timeout = max(600, getOption("timeout")))
  
  # Download the file
  download.file(url, destfile = destfile, method = "auto", mode = "wb")
  
  # Load the .nc file into a raster object
  assign(paste0("tmin8.5_", year_range), terra::rast(destfile))
  
  # Print confirmation of download and extraction
  print(paste("Downloaded and loaded:", destfile))
}



## 1.3. tmax DATA
# 4.5, tmax: 2006-2099, monthly
# urls_tmax_45_1 <- c(
#   "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1/macav2livneh_tasmax_bcc-csm1-1_r1i1p1_rcp45_2006_2025_CONUS_monthly.nc",
#   "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1/macav2livneh_tasmax_bcc-csm1-1_r1i1p1_rcp45_2026_2045_CONUS_monthly.nc",
#   "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1/macav2livneh_tasmax_bcc-csm1-1_r1i1p1_rcp45_2046_2065_CONUS_monthly.nc",
#   "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1/macav2livneh_tasmax_bcc-csm1-1_r1i1p1_rcp45_2066_2085_CONUS_monthly.nc",
#   "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1/macav2livneh_tasmax_bcc-csm1-1_r1i1p1_rcp45_2086_2099_CONUS_monthly.nc"
# )
# 
# 
# # 8.5, tmax: 2006-2099, monthly
# urls_tmax_85_1 <- c(
#   "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1/macav2livneh_tasmax_bcc-csm1-1_r1i1p1_rcp85_2006_2025_CONUS_monthly.nc",
#   "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1/macav2livneh_tasmax_bcc-csm1-1_r1i1p1_rcp85_2026_2045_CONUS_monthly.nc",
#   "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1/macav2livneh_tasmax_bcc-csm1-1_r1i1p1_rcp85_2046_2065_CONUS_monthly.nc",
#   "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1/macav2livneh_tasmax_bcc-csm1-1_r1i1p1_rcp85_2066_2085_CONUS_monthly.nc",
#   "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1/macav2livneh_tasmax_bcc-csm1-1_r1i1p1_rcp85_2086_2099_CONUS_monthly.nc"
# )

urls_tmax_45_1 <- c(
  "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1-m/macav2livneh_tasmax_bcc-csm1-1-m_r1i1p1_rcp45_2006_2025_CONUS_monthly.nc",
  "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1-m/macav2livneh_tasmax_bcc-csm1-1-m_r1i1p1_rcp45_2026_2045_CONUS_monthly.nc",
  "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1-m/macav2livneh_tasmax_bcc-csm1-1-m_r1i1p1_rcp45_2046_2065_CONUS_monthly.nc",
  "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1-m/macav2livneh_tasmax_bcc-csm1-1-m_r1i1p1_rcp45_2066_2085_CONUS_monthly.nc",
  "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1-m/macav2livneh_tasmax_bcc-csm1-1-m_r1i1p1_rcp45_2086_2099_CONUS_monthly.nc"
)


# 8.5, tmax: 2006-2099, monthly
urls_tmax_85_1 <- c(
  "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1-m/macav2livneh_tasmax_bcc-csm1-1-m_r1i1p1_rcp85_2006_2025_CONUS_monthly.nc",
  "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1-m/macav2livneh_tasmax_bcc-csm1-1-m_r1i1p1_rcp85_2026_2045_CONUS_monthly.nc",
  "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1-m/macav2livneh_tasmax_bcc-csm1-1-m_r1i1p1_rcp85_2046_2065_CONUS_monthly.nc",
  "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1-m/macav2livneh_tasmax_bcc-csm1-1-m_r1i1p1_rcp85_2066_2085_CONUS_monthly.nc",
  "http://thredds.northwestknowledge.net:8080/thredds/fileServer/NWCSC_INTEGRATED_SCENARIOS_ALL_CLIMATE/macav2livneh/bcc-csm1-1-m/macav2livneh_tasmax_bcc-csm1-1-m_r1i1p1_rcp85_2086_2099_CONUS_monthly.nc"
)

# Loop to download and extract each file
for (url in urls_tmax_45_1) {
  # Extract year range from the URL for naming the file
  year_range <- gsub(".*_([0-9]{4}_[0-9]{4})_.*", "\\1", url)
  
  # Set destination file name
  destfile <- paste0("/Volumes/Extreme SSD/UNL/DWFI/projection/data/2/tmax4.5_", year_range, ".nc")
  
  # Increase timeout in case of large files
  options(timeout = max(600, getOption("timeout")))
  
  # Download the file
  download.file(url, destfile = destfile, method = "auto", mode = "wb")
  
  # Load the .nc file into a raster object
  assign(paste0("tmax4.5_", year_range), terra::rast(destfile))
  
  # Print confirmation of download and extraction
  print(paste("Downloaded and loaded:", destfile))
}


# Loop to download and extract each file
for (url in urls_tmax_85_1) {
  # Extract year range from the URL for naming the file
  year_range <- gsub(".*_([0-9]{4}_[0-9]{4})_.*", "\\1", url)
  
  # Set destination file name
  destfile <- paste0("/Volumes/Extreme SSD/UNL/DWFI/projection/data/2/tmax8.5_", year_range, ".nc")
  
  # Increase timeout in case of large files
  options(timeout = max(600, getOption("timeout")))
  
  # Download the file
  download.file(url, destfile = destfile, method = "auto", mode = "wb")
  
  # Load the .nc file into a raster object
  assign(paste0("tmax8.5_", year_range), terra::rast(destfile))
  
  # Print confirmation of download and extraction
  print(paste("Downloaded and loaded:", destfile))
}

## 2. Estimating Variables
temp_base <- 12  # base temperature for GDD calculation
temp_up <- 31  # threshold temperature for EDD calculation

# Load US counties
US_counties <- tigris::counties(state = "US", cb = TRUE) %>%
  st_as_sf()

data_dir <- "/Volumes/Extreme SSD/UNL/DWFI/projection/data/2"

# Define file paths for 4.5 and 8.5 scenarios
file_paths <- list(
  `4.5` = list(
    pr = list.files(path = data_dir, pattern = "pr4.5.*\\.nc$", full.names = TRUE),
    tmin = list.files(path = data_dir, pattern = "tmin4.5.*\\.nc$", full.names = TRUE),
    tmax = list.files(path = data_dir, pattern = "tmax4.5.*\\.nc$", full.names = TRUE)
  ),
  `8.5` = list(
    pr = list.files(path = data_dir, pattern = "pr8.5.*\\.nc$", full.names = TRUE),
    tmin = list.files(path = data_dir, pattern = "tmin8.5.*\\.nc$", full.names = TRUE),
    tmax = list.files(path = data_dir, pattern = "tmax8.5.*\\.nc$", full.names = TRUE)
  )
)

# Function to process one month of data
process_monthly_data <- function(year, scenario, file_paths, US_counties) {
  # Load corresponding files for the scenario and year range
  pr_file <- file_paths[[scenario]]$pr[grepl(sprintf("%d", year), file_paths[[scenario]]$pr)]
  tmin_file <- file_paths[[scenario]]$tmin[grepl(sprintf("%d", year), file_paths[[scenario]]$tmin)]
  tmax_file <- file_paths[[scenario]]$tmax[grepl(sprintf("%d", year), file_paths[[scenario]]$tmax)]
  
  # Check if all files exist
  if (length(pr_file) == 0 || length(tmin_file) == 0 || length(tmax_file) == 0) {
    message(sprintf("Skipping %d-%02d for %s scenario due to missing files.", year, month, scenario))
    return(NULL)
  }
  
  # Load raster data
  pr_raster <- rast(pr_file)
  tmin_raster <- rast(tmin_file)
  tmax_raster <- rast(tmax_file)
  
  names(pr_raster) <- time(pr_raster)
  names(tmin_raster) <- time(tmin_raster)
  names(tmax_raster) <- time(tmax_raster)
  
  # Assign CRS to raster and vector dataset
  US_counties_sf <- st_transform(US_counties, crs(pr_raster))
  
  # Extract data for each variable
  pr_data <- as.data.table(terra::extract(pr_raster, vect(US_counties_sf)))
  tmin_data <- as.data.table(terra::extract(tmin_raster, vect(US_counties_sf)))
  tmax_data <- as.data.table(terra::extract(tmax_raster, vect(US_counties_sf)))
  
  # Reshape and aggregate data
  pr_data_ID <- melt(pr_data, id.vars = "ID", variable.name = "Date", value.name = "ppt") %>%
    .[, .(ppt = mean(ppt, na.rm = TRUE)), by = .(ID, Date)]
  
  tmin_data_ID <- melt(tmin_data, id.vars = "ID", variable.name = "Date", value.name = "tmin") %>%
    .[, .(tmin = mean(tmin, na.rm = TRUE)), by = .(ID, Date)]
  
  tmax_data_ID <- melt(tmax_data, id.vars = "ID", variable.name = "Date", value.name = "tmax") %>%
    .[, .(tmax = mean(tmax, na.rm = TRUE)), by = .(ID, Date)]
  
  setkey(pr_data_ID, ID, Date)
  setkey(tmin_data_ID, ID, Date)
  setkey(tmax_data_ID, ID, Date)
  
  
  # Merge the data
  weather_data <- merge(pr_data_ID, tmin_data_ID, by = c("ID", "Date"), all = TRUE)
  weather_data <- merge(weather_data, tmax_data_ID, by = c("ID", "Date"), all = TRUE)
  
  # Add year and month
  weather_data[, year := year(Date)]
  weather_data[, month := month(Date)]
  
  US_counties_v_1 <- st_as_sf(vect(US_counties_sf)) %>%
    mutate(ID = seq_len(nrow(.))) %>%
    as.data.table()
  
  US_counties_v_1 <- merge(US_counties_v_1, weather_data, by = "ID", all.x = TRUE)
  
  US_counties_v_1[, tmax := tmax - 273.15]
  US_counties_v_1[, tmin := tmin - 273.15]
  
  US_counties_v_1[, Tavg := (tmax + tmin) / 2]
  US_counties_v_1[, month_heat_index := (Tavg / 5) ^ 1.514]
  
  # US_counties_v_2 <- US_counties_v_1[, .(annual_heat_index = sum(month_heat_index, na.rm = TRUE)), by = .(year, NAME, STATE_NAME)]
  
  US_counties_annual <- US_counties_v_1[, .(
    annual_heat_index = sum(month_heat_index, na.rm = TRUE)
  ), by = .(year, NAME, STATE_NAME)]
  
  US_counties_v_2 <-  merge(US_counties_v_1, US_counties_annual, by = c("year", "NAME", "STATE_NAME"), all.x = TRUE)
  
  US_counties_v_2[, a := 675 * 10^(-9) * annual_heat_index^3 - 771 * 10^(-7) * annual_heat_index^2 + 1792 * 10^(-5) * annual_heat_index + 0.49239]
  
  # joined_weather_SPEI <- merge(US_counties_v_1, US_counties_v_2, by = c("year", "NAME", "STATE_NAME"), all.x = TRUE)
  US_counties_v_2[, PET := ifelse(Tavg <= 0, 0, 16 * ((10 * Tavg) / annual_heat_index) ^ a)]
  US_counties_v_2[, D := ppt - PET]
  
  # joined_weather_SPEI[, Tavg := (tmax + tmin) / 2]
  US_counties_v_2[, GDD := fifelse(Tavg < temp_base, 0,
                                   fifelse(Tavg > temp_up, temp_up - temp_base,
                                           Tavg - temp_base))]
  
  US_counties_v_2[, EDD := fifelse(tmax > temp_up, tmax - temp_up, 0)]
  
  print(sprintf("Done with year: %d", year))
  
  joined_weather <- US_counties_v_2[, .(
    avg_tmin = tmin,
    avg_tmax = tmax,
    avg_tmean = Tavg,
    sum_ppt = sum(ppt, na.rm = TRUE),
    monthly_D = sum(D, na.rm = TRUE),
    monthly_GDD = sum(GDD, na.rm = TRUE),
    monthly_EDD = sum(EDD, na.rm = TRUE)
  ), by = .(year, month, NAME, STATE_NAME)]
  
  return(joined_weather)
}

# 2006, 2026, 2046, 2066, 2086
month_year_data <- CJ(year = 2086)

# Process all data
results <- list()
for (scenario in names(file_paths)) {
  scenario_results <- lapply(1:nrow(month_year_data), function(i) {
    year <- month_year_data[i, year]
    process_monthly_data(year, scenario, file_paths, US_counties)
  })
  results[[scenario]] <- do.call(rbind, scenario_results)
}

# Save results
saveRDS(results$`4.5`, file.path(data_dir, "2086_scenario_4.5_results.rds"))
saveRDS(results$`8.5`, file.path(data_dir, "2086_scenario_8.5_results.rds"))

## 3. Combining datasets  

# 4.5
drought_2006_45 <- readRDS("/Volumes/Extreme SSD/UNL/DWFI/projection/data/2/2006_scenario_4.5_results.rds")
drought_2026_45 <- readRDS("/Volumes/Extreme SSD/UNL/DWFI/projection/data/2/2026_scenario_4.5_results.rds")
drought_2046_45 <- readRDS("/Volumes/Extreme SSD/UNL/DWFI/projection/data/2/2046_scenario_4.5_results.rds")
drought_2066_45 <- readRDS("/Volumes/Extreme SSD/UNL/DWFI/projection/data/2/2066_scenario_4.5_results.rds")
drought_2086_45 <- readRDS("/Volumes/Extreme SSD/UNL/DWFI/projection/data/2/2086_scenario_4.5_results.rds")

drought_data_4.5 <- rbind(drought_2006_45, drought_2026_45,
                          drought_2046_45, drought_2066_45, drought_2086_45)
saveRDS(drought_data_4.5, file.path(data_dir, "drought_4.5_ALL.rds"))


# 8.5
drought_2006_85 <- readRDS("/Volumes/Extreme SSD/UNL/DWFI/projection/data/2/2006_scenario_8.5_results.rds")
drought_2026_85 <- readRDS("/Volumes/Extreme SSD/UNL/DWFI/projection/data/2/2026_scenario_8.5_results.rds")
drought_2046_85 <- readRDS("/Volumes/Extreme SSD/UNL/DWFI/projection/data/2/2046_scenario_8.5_results.rds")
drought_2066_85 <- readRDS("/Volumes/Extreme SSD/UNL/DWFI/projection/data/2/2066_scenario_8.5_results.rds")
drought_2086_85 <- readRDS("/Volumes/Extreme SSD/UNL/DWFI/projection/data/2/2086_scenario_8.5_results.rds")

drought_data_8.5 <- rbind(drought_2006_85, drought_2026_85,
                          drought_2046_85, drought_2066_85, drought_2086_85)
saveRDS(drought_data_8.5, file.path(data_dir, "drought_8.5_ALL.rds"))


```
