---
title: "Tables and Figures"
format: html
editor: visual
---

### ==============================================

### Visualization and Results

### This script generates figures and tables for the paper:

### Figure 1. Annual US national beef cow inventory and pastureland data.

### Figure 2. Projected cattle stocking rates (2023-2099).

### Figure 3. County level SR for 2023/2090 (map).

### Appendix C figures (C1, C2, C3)

### Table 2. Stocking rate by US geographical regions, 1982-2022s.

### Table 3. Summary table of all variables

### Table 4. Regression analysis results for delayed drought effects.

### Table 5. PDSI and USDM regression results 

### Table B1. Correlation matrix for variables used in the stocking rate regression model.

### ===============================================


### Figure 1. Annual US national beef cow inventory and pastureland data.

```{r}

# Create ANNUAL pastureland and cattle inventory data - ALL US
merged_ALL_national <- merged_ALL %>% 
  group_by(year) %>%
  summarise(
    pastureland_all_types = sum(pastureland_all_types, na.rm = TRUE),
     beef_inv = sum(beef_inv, na.rm = TRUE)
  )

merged_ALL_national$beef_inv <- merged_ALL_national$beef_inv/1000000
merged_ALL_national$pastureland_all_types <- merged_ALL_national$pastureland_all_types/1000000
scaling_factor <- max(merged_ALL_national$pastureland_all_types) / max(merged_ALL_national$beef_inv)

merged_ALL_national <- merged_ALL_national %>% 
  filter(year != 1978)

ggplot() +
  geom_line(data = merged_ALL_national, aes(x = year, y = beef_inv, color = "Beef cow inventory, head")) +  # Assign legend label
  geom_line(data = merged_ALL_national, aes(x = year, y = pastureland_all_types / scaling_factor, color = "Pastureland (all), acres")) +  # Assign legend label and scaling factor
  xlab("Year") + 
  theme_bw() + 
  labs(title = "") + 
  # ylab("Steer, live finishing weight (lbs)") +
  scale_x_continuous(breaks = seq(min(merged_ALL_national$year), max(merged_ALL_national$year), by = 5)) +  # Show year as numeric on x-axis
  scale_y_continuous(
    name = "Beef cow inventory, head (mln)",
    sec.axis = sec_axis(~ . * scaling_factor, name = "Pastureland (all), acres (mln)")  # Multiply back the scaled beef_head for the secondary y-axis
  ) +
  scale_color_manual(values = c("Beef cow inventory, head" = "black", "Pastureland (all), acres" = "red"),  # Manually assign colors
                     name = "") +  # Legend title
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 10),
        axis.title.y = element_text(size = 10),
        axis.title.y.right = element_text(size = 10),  
        axis.text.y = element_text(size = 10),
        legend.position = c(0.8, 0.8),  # Position the legend inside the plot at coordinates (0.2, 0.8)
        legend.background = element_rect(fill = alpha('white', 0.5)),
        legend.text = element_text(size = 12),  # Set legend text size
        legend.title = element_text(size = 14))   # Make legend background semi-transparent

```

### Figure 2. Projected cattle stocking rates (2020-2099).

```{r}

# FIGURE 2
ggplot(average_stock_rate_US, 
       aes(x = year, y = mean_projected_stock_rate, color = scenario, fill = scenario)) +
  
  geom_ribbon(
    aes(ymin = ci_lower, ymax = ci_upper),
    alpha = 0.2,
    color = NA
  ) +
  
  geom_smooth(
    se = FALSE,
    method = "loess",
    span = 0.35,
    linewidth = 1.5
  ) +
  
  labs(
    x = "Year",
    y = "Projected Stock Rate",
    color = "Scenario",
    fill = "Scenario"
  ) +
  
  scale_color_manual(values = c("RCP4.5" = "blue", "RCP8.5" = "red")) +
  scale_fill_manual(values = c("RCP4.5" = "blue", "RCP8.5" = "red")) +
  
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 10)
  )
   

```


### Figure 3. County level SR for 2022/2099 (map).

```{r}

# FIGURE 3
options(tigris_use_cache = TRUE)
map_input <- combined_data %>%
  filter(year %in% c(2023, 2090)) %>%
  group_by(county, state, year, scenario) %>%
  summarise(
    SR_hat = mean(SR_hat, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  # Clean scenario labels for facet strips
  mutate(
    scenario = recode(scenario,
      "RCP 4.5" = "RCP 4.5",
      "RCP 8.5" = "RCP 8.5"
    )
  )

us_counties <- counties(cb = TRUE, year = 2022, class = "sf") %>%
  mutate(
    state  = tolower(STATE_NAME),
    county = tolower(NAME)
  )
us_states <- states(cb = TRUE, year = 2022, class = "sf")

map_data <- us_counties %>%
  left_join(map_input, by = c("state", "county")) %>%
  filter(!is.na(year), !is.na(scenario))

map_data$stock_bin <- cut(
  map_data$SR_hat,
  breaks = c(0, 0.01, 0.02, 0.03, 0.04, 0.06, 0.08, 0.11, Inf),
  labels = c("0–0.01", "0.011–0.02", "0.021–0.03", "0.031–0.04", "0.041–0.06",
             "0.061–0.08", "0.081–0.10", ">0.11"),
  include.lowest = TRUE,
  right = TRUE
)

# Color scale: light = low stocking (arid West), dark = high stocking (SE)
sr_colors <- c(
  "0–0.01"    = "#4d0000",   # very light cream
  "0.011–0.02" = "#b30000",   # light orange
  "0.021–0.03" = "red",   # light orange
  "0.031–0.04" = "#e34a33",   # light orange
  "0.041–0.06" = "#fc8d59",   # orange        — modal range
  "0.061–0.08" = "#fee08b",   # orange-red
  "0.081–0.10" = "yellow",   # red
  ">0.11"    = "green"    # very dark red  — high stocking SE/Appalachia
)

p_levels <- ggplot(map_data) +
  geom_sf(aes(fill = stock_bin), color = NA) +
  geom_sf(data = us_states, fill = NA, color = "black",
          linewidth = 0.15) +
  scale_fill_manual(
    values   = sr_colors,
    name     = "Stocking Rate\n(AUY/acre)",
    drop     = FALSE,
    na.value = "gray85"
  ) +
  facet_grid(year ~ scenario) +
  coord_sf(
    xlim   = c(-125, -66.5),
    ylim   = c(24, 50),
    expand = FALSE
  ) +
  theme_minimal(base_size = 16) +
  theme(
    axis.text        = element_blank(),
    axis.ticks       = element_blank(),
    panel.grid       = element_blank(),
    axis.title       = element_blank(),
    strip.text       = element_text(size = 16, face = "bold"),
    legend.title     = element_text(size = 16, face = "bold"),
    legend.text      = element_text(size = 16),
    legend.position  = "right",
    legend.key.width = unit(0.8, "cm"),
    plot.margin      = margin(5, 5, 5, 5)
  )


pct_data <- map_input %>%
  pivot_wider(
    names_from  = year,
    values_from = SR_hat,
    names_prefix = "yr_"
  ) %>%
  mutate(
    perc_change = 100 * (yr_2099 - yr_2023) / yr_2023
  ) %>%
  filter(!is.na(perc_change))

map_pct <- us_counties %>%
  left_join(pct_data, by = c("state", "county")) %>%
  filter(!is.na(scenario))

map_pct$perc_bin <- cut(
  map_pct$perc_change,
  breaks = c(-Inf, -50, -25, 0, 25, Inf),
  labels = c("<-50%", "-50% to -25%", "-25% to 0%",
             "0% to 25%", ">25%"),
  include.lowest = TRUE,
  right = FALSE
)

pct_colors <- c(
  "<-50%"        = "#67000d",   # very dark red
  "-50% to -25%" = "#cb181d",   # red
  "-25% to 0%"   = "#fb6a4a",   # light red-orange
  "0% to 25%"    = "#fdcc8a",
  ">25%" = "green"
)

p_change <- ggplot(map_pct) +
  geom_sf(aes(fill = perc_bin), color = NA) +
  geom_sf(data = us_states, fill = NA, color = "black",
          linewidth = 0.25) +
  scale_fill_manual(
    values   = pct_colors,
    name     = "% Change in\nStocking Rate",
    drop     = FALSE,
    na.value = "gray85"
  ) +
  facet_wrap(~ scenario, nrow = 1) +
  coord_sf(
    xlim   = c(-125, -66.5),
    ylim   = c(24, 50),
    expand = FALSE
  ) +
  theme_minimal(base_size = 16) +
  theme(
    axis.text        = element_blank(),
    axis.ticks       = element_blank(),
    panel.grid       = element_blank(),
    axis.title       = element_blank(),
    strip.text       = element_text(size = 16, face = "bold"),
    legend.title     = element_text(size = 16, face = "bold"),
    legend.text      = element_text(size = 16),
    legend.position  = "right",
    legend.key.width = unit(0.8, "cm"),
    plot.margin      = margin(5, 5, 5, 5)
  )

p_levels
p_change

```


### Appendix C figures (C1, C2, C3)

```{r}

# C1
p_GDD <- ggplot(all_r_full, aes(x = year, y = mean_GDD, color = scenario)) +
  geom_line(linewidth = 0.8, alpha = 0.6) +
  geom_smooth(se = FALSE, method = "loess", span = 0.3, linewidth = 1.5) +
  scale_color_manual(values = c("RCP4.5" = "steelblue", "RCP8.5" = "red3")) +
  labs(
    title    = "Growing Degree Days (GDD) by Scenario",
    subtitle = "Higher GDD = more beneficial heat accumulation = higher stocking rate",
    x = "Year", y = "Mean GDD (degree-days)",
    color = "Scenario",
    caption = paste0(
      "GDD base = 15°C, cap = 31°C. Averaged across 10 realizations.\n",
      "RCP 8.5 higher GDD early explains near-term SR advantage.\n",
      "RCP 4.5 GDD plateaus mid-century — explains SR stabilization."
    )
  ) +
  theme_classic(base_size = 12) +
  theme(plot.caption = element_text(size = 8, color = "gray40"))
print(p_GDD)


# C2
p_EDD <- ggplot(all_r_full, aes(x = year, y = mean_EDD, color = scenario)) +
  geom_line(linewidth = 0.8, alpha = 0.6) +
  geom_smooth(se = FALSE, method = "loess", span = 0.3, linewidth = 1.5) +
  scale_color_manual(values = c("RCP4.5" = "steelblue", "RCP8.5" = "red3")) +
  labs(
    title    = "Extreme Degree Days (EDD) by Scenario",
    subtitle = "Higher EDD = more damaging heat = lower stocking rate",
    x = "Year", y = "Mean EDD (degree-days above 31°C)",
    color = "Scenario",
    caption = paste0(
      "EDD threshold = 31°C (empirically estimated from hay yield analysis).\n",
      "RCP 4.5 EDD stabilizes after ~2050 — explains why SR plateaus.\n",
      "RCP 8.5 EDD accelerates — explains persistent SR decline."
    )
  ) +
  theme_classic(base_size = 12) +
  theme(plot.caption = element_text(size = 8, color = "gray40"))
print(p_EDD)



# C3
p_D <- ggplot(all_r_full, aes(x = year, y = mean_D, color = scenario)) +
  geom_line(linewidth = 0.8, alpha = 0.6) +
  geom_smooth(se = FALSE, method = "loess", span = 0.3, linewidth = 1.5) +
  scale_color_manual(values = c("RCP4.5" = "steelblue", "RCP8.5" = "red3")) +
  labs(
    title    = "Water Deficit (D) by Scenario",
    subtitle = "Higher D = greater drought stress = lower stocking rate",
    x = "Year", y = "Mean Water Deficit (mm)",
    color = "Scenario",
    caption = "Positive D = moisture deficit. Averaged across 10 realizations."
  ) +
  theme_classic(base_size = 12) +
  theme(plot.caption = element_text(size = 8, color = "gray40"))
print(p_D)

```


### Table 2. Stocking rate by US geographical regions, 1982-2022s.

```{r}

# TABLE 2
merged_ALL_regional_SR <- regional_SR %>%
  filter(is.finite(stock_rate_AUY_regional)) %>%   # fixed: correct column name
  mutate(decade = floor(year / 10) * 10)

summary_table <- merged_ALL_regional_SR %>%
  group_by(region, decade) %>%
  summarize(
    avg_stock_rate = weighted.mean(stock_rate_AUY_regional,    # fixed: correct column
                                   w    = total_pastureland,   # weight by pastureland size
                                   na.rm = TRUE),
    .groups = "drop"
  )

summary_table_wide <- summary_table %>%
  pivot_wider(
    names_from  = decade,
    values_from = avg_stock_rate,
    names_prefix = "Decade_"
  ) %>%
  mutate(
    change_1990_1980 = ((Decade_1990 - Decade_1980) / Decade_1980) * 100,
    change_2000_1990 = ((Decade_2000 - Decade_1990) / Decade_1990) * 100,
    change_2010_2000 = ((Decade_2010 - Decade_2000) / Decade_2000) * 100,
    change_2020_2010 = ((Decade_2020 - Decade_2010) / Decade_2010) * 100
  )

# National average — aggregate from totals, not average of regional rates
national_avg <- merged_ALL_regional_SR %>%
  group_by(decade) %>%
  summarize(
    avg_stock_rate = sum(total_AUs * (avg_grazing_months / 12), na.rm = TRUE) /
                     sum(total_pastureland, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  pivot_wider(
    names_from  = decade,
    values_from = avg_stock_rate,
    names_prefix = "Decade_"
  ) %>%
  mutate(
    region           = "National Avg",
    change_1990_1980 = ((Decade_1990 - Decade_1980) / Decade_1980) * 100,
    change_2000_1990 = ((Decade_2000 - Decade_1990) / Decade_1990) * 100,
    change_2010_2000 = ((Decade_2010 - Decade_2000) / Decade_2000) * 100,
    change_2020_2010 = ((Decade_2020 - Decade_2010) / Decade_2010) * 100
  )

final_table <- bind_rows(summary_table_wide, national_avg) %>%
  mutate(across(where(is.numeric), ~ round(., 4)))

ft <- flextable(data = final_table)
ft

```

### Table 3. Summary table of all variables

```{r}

# TABLE 3
SR_data <- merged_ALL_SR %>%
  filter(
    is.finite(stock_rate_AUY),
    stock_rate_AUY <= 10
  )

summary_data <- left_join(drought_final, hay_all_stocks_1000t, 
                           by = c("year", "state")) %>%
  left_join(SR_data, by = c("year", "state", "county")) %>%
  na.omit()

table3 <- summary_data %>%
  ungroup() %>%
  summarise(
    across(
      c(stock_rate_AUY, total_D, total_GDD, total_EDD, hay_stock_1000t),
      list(
        Mean    = ~mean(.x, na.rm = TRUE),
        Std.Dev = ~sd(.x,   na.rm = TRUE),
        Min     = ~min(.x,  na.rm = TRUE),
        Max     = ~max(.x,  na.rm = TRUE)
      ),
      .names = "{.col}_{.fn}"
    ),
    Obs = n()
  )

table3
```

### Table 4. Regression analysis results for delayed drought effects.

```{r}

# TABLE 4
county_fe <- "county"
year_fe   <- "year"
state_fe <- "state"
region_fe <- "region"
FE        <- paste(county_fe, year_fe, sep = " + ")  
# Options:
# FE <- "county"            # County FE only
# FE <- "year"              # Year FE only  
# FE <- "county + year"     # County + Year FE (preferred)

model_formulas <- list(
  m1 = as.formula(paste("log(stock_rate_AUY) ~ total_D |", FE)),
  m2 = as.formula(paste("log(stock_rate_AUY) ~ total_D + drought_lag1 |", FE)),
  m3 = as.formula(paste("log(stock_rate_AUY) ~ total_D + drought_lag1 + total_GDD |", FE)),
  m4 = as.formula(paste("log(stock_rate_AUY) ~ total_D + drought_lag1 + total_GDD + total_EDD |", FE)),
  m5 = as.formula(paste("log(stock_rate_AUY) ~ total_D + drought_lag1 + total_GDD + total_EDD + hay_stock_1000t_lagged |", FE)),
  m6 = as.formula(paste("log(stock_rate_AUY) ~ total_D + drought_lag1 + total_GDD + total_EDD + neighbor_avg_hay_stock_1000t_lagged |", FE))
)

model_data <- list(
  m1 = regression_data_lagged,
  m2 = regression_data_lagged,
  m3 = regression_data_lagged,
  m4 = regression_data_lagged,
  m5 = regression_data_lagged,
  m6 = regression_data_lagged_neighbour
)

regs <- mapply(
  FUN      = function(formula, data) {
    feols(formula, data = data, cluster = ~county)
  },
  formula  = model_formulas,
  data     = model_data,
  SIMPLIFY = FALSE
)

do.call(etable, c(
  regs,
  list(
    digits   = 4,
    se.below = TRUE,
    signif.code = c("***" = 0.01, "**" = 0.05, "*" = 0.1),
    headers  = list("(1)" = 1, "(2)" = 1, "(3)" = 1,
                    "(4)" = 1, "(5)" = 1, "(6)" = 1)
  )
))


```


### Table 5. PDSI and USDM regression results 

```{r}

# TABLE 5
PDSI_monthly <- read.csv("/Users/sharaakat/Dropbox/akat_shara/DWFI_drought/PDSI_1981_2022_monthly.csv")
region_grazing_months <- data.frame(
  region      = c("Appalachia", "Corn Belt", "Delta States", "Lake States",
                  "Mountain", "Northeast", "Northern Plains",
                  "Pacific", "Southeast", "Southern Plains"),
  start_month = c(3, 4, 4, 4, 4, 4, 4, 3, 2, 4),
  end_month   = c(10, 10, 10, 9, 9, 9, 10, 10, 10, 10)
)


PDSI_grazing <- PDSI_monthly %>%
  mutate(state = tolower(state), county = tolower(county)) %>%
  mutate(region = state_to_region(state)) %>%
  left_join(region_grazing_months, by = "region") %>%
  filter(!is.na(region),
         month >= start_month,
         month <= end_month)


PDSI_annual <- PDSI_grazing %>%
  group_by(state, county, year) %>%
  summarise(
    pdsi_mean = mean(index, na.rm = TRUE),   # mean PDSI over grazing season
    .groups   = "drop"
  )


# Dataset A: USDM merge
USDM_reg <- regression_data_lagged %>%
  left_join(
    USDM_annual %>% mutate(state = tolower(state), county = tolower(county)),
    by = c("state", "county", "year")
  ) %>%
  filter(year >= 2002, year <= 2022) %>%
  filter(!is.na(D0_weeks), !is.na(stock_rate_AUY))


PDSI_reg <- regression_data_lagged %>%
  left_join(PDSI_annual, by = c("state", "county", "year")) %>%
  filter(year >= 1981, year <= 2022) %>%
  filter(!is.na(pdsi_mean), !is.na(stock_rate_AUY))

m_usdm1 <- feols(
  log(stock_rate_AUY) ~
    D0_weeks + D1_weeks + D2_weeks + D3_weeks + D4_weeks |
    county + year,
  data    = USDM_reg,
  cluster = ~county
)

m_usdm2 <- feols(
  log(stock_rate_AUY) ~
    D0_weeks + D1_weeks + D2_weeks + D3_weeks + D4_weeks +
    hay_stock_1000t_lagged |
    county + year,
  data    = USDM_reg,
  cluster = ~county
)

m_usdm3 <- feols(
  log(stock_rate_AUY) ~
    D0_weeks + D1_weeks + D2_weeks + D3_weeks + D4_weeks +
    hay_stock_1000t_lagged +
    total_GDD + total_EDD |     # NO total_D — USDM already captures drought
    county + year,
  data    = USDM_reg,
  cluster = ~county
)

# Col 4: Main model spec on USDM sample — direct apples-to-apples comparison
m_weather_usdmsample <- feols(
  log(stock_rate_AUY) ~
    total_D + drought_lag1 + total_GDD + total_EDD +
    hay_stock_1000t_lagged |
    county + year,
  data    = USDM_reg,
  cluster = ~county
)


m_pdsi1 <- feols(
  log(stock_rate_AUY) ~
    pdsi_mean |
    county + year,
  data    = PDSI_reg,
  cluster = ~county
)

m_pdsi2 <- feols(
  log(stock_rate_AUY) ~
    pdsi_mean +
    hay_stock_1000t_lagged |
    county + year,
  data    = PDSI_reg,
  cluster = ~county
)

m_pdsi3 <- feols(
  log(stock_rate_AUY) ~
    pdsi_mean +
    hay_stock_1000t_lagged +
    total_GDD + total_EDD |
    county + year,
  data    = PDSI_reg,
  cluster = ~county
)


etable(
  list(m_usdm1, m_usdm2, m_usdm3, m_weather_usdmsample,
       m_pdsi1, m_pdsi2, m_pdsi3),
  se          = "cluster",
  signif.code = c("***" = 0.01, "**" = 0.05, "*" = 0.1),
  keep_raw    = c("D0_weeks", "D1_weeks", "D2_weeks", "D3_weeks", "D4_weeks",
                  "hay_stock_1000t_lagged", "total_D", "drought_lag1",
                  "total_GDD", "total_EDD", "pdsi_mean"),
  dict        = c(
    D0_weeks               = "# weeks in D0",
    D1_weeks               = "# weeks in D1",
    D2_weeks               = "# weeks in D2",
    D3_weeks               = "# weeks in D3",
    D4_weeks               = "# weeks in D4",
    hay_stock_1000t_lagged = "Hay stock_{s,t-1}",
    total_D                = "D_{ist}",
    drought_lag1           = "D_{is,t-1}",
    total_GDD              = "GDD_{ist}",
    total_EDD              = "EDD_{ist}",
    pdsi_mean              = "PDSI (grazing season mean)"
  ),
  fitstat     = ~ r2 + wr2 + n
)


```


### Table B1. Correlation matrix for variables used in the stocking rate regression model.

```{r}

# TABLE A1
corr_vars <- regression_data_lagged %>%
  dplyr::select(
    total_D, drought_lag1,
    total_GDD, total_EDD,
    hay_stock_1000t
  ) %>%
  na.omit()

# Compute Pearson correlation matrix
corr_matrix <- round(cor(corr_vars, method = "pearson"), 2)

# Print matrix in console
View(corr_matrix)
print(corr_matrix)


```


